---
title: "DE_analyses and WGCNA"
author: "Colette Berg"
date: "2025-10-07"
output: html_document
---

# set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(edgeR)
library(DESeq2)
library(genefilter)
library(tidyverse)
library(WGCNA)
library(clusterProfiler)
library(GO.db)
library(reshape2)
library(fs)

setwd("~/Dropbox/LowryLab/data")
```

# read in the gff3
```{r}
# get number of genes in genome 
AP13_gff <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/AP13/Pvirgatum_AP13_noHeader.gff3", sep = "\t")

# getting list of genes only
AP13_geneList <- filter(AP13_gff, gene == "gene")

```

# process the raw data from HTSeq
```{r read in data}
# read in read counts from HTSeq; 
# not including the F1's 

# get a list of all the gene counts files 
tsv_files <- dir_ls(path = "~/Dropbox/LowryLab/data/expression/HTSeq_Counts_pseudo/", regexp = "\\.tsv$")

# read all of them into one file
all_data <- read_tsv(tsv_files, id = "file_source", col_names = FALSE) # The 'id' argument adds a column with the filename

# clean up the individual ID column
all_data$file_source <- gsub("/Users/coletteberg/Dropbox/LowryLab/expression/data/HTSeq_Counts_pseudo/", "", all_data$file_source)

all_data$file_source <- gsub(".geneID.counts.tsv", "", all_data$file_source)

# clean up the gene ID column
all_data$X1 <- gsub(".v6.1", "", all_data$X1)

# now add a genome column by string-splitting
allData_sep <- separate_wider_delim(all_data, X1, delim="_", names=c("Genome", "Gene"), too_many = "debug")
allData_clean <- filter(allData_sep, X1_pieces == 2)

# split by genome
AP13_counts <- filter(dplyr::select(allData_clean, -X1_ok, -X1_pieces, -X1_remainder, -X1), Genome == "AP13")
DAC6_counts <- filter(dplyr::select(allData_clean, -X1_ok, -X1_pieces, -X1_remainder, -X1), Genome == "DAC6")

# clean up count column
AP13_counts  <- rename(AP13_counts, AP13_count = X2)
DAC6_counts  <- rename(DAC6_counts, DAC6_count = X2)

# paste together
both_counts <- cbind(AP13_counts, dplyr::select(DAC6_counts, DAC6_count))

# sum up the counts
both_counts <- mutate(both_counts, counts = AP13_count + DAC6_count)

# keep the necessary columns
counts_clean_long <- dplyr::select(both_counts, file_source, Gene, counts)

# clean up name 
counts_clean_long <- rename(counts_clean_long, indv = file_source)

# pivot wider 
counts_wide <- pivot_wider(counts_clean_long, names_from=indv, values_from=counts)

# read in metadata 
metaData <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/expression/Pvirgatum.fourWay.RNAseq.libraryInfo.txt", sep="\t")

# format wide dataframe for filtering
counts_wide <- as.data.frame(counts_wide)
rownames(counts_wide) <- counts_wide$Gene

# transpose
counts_t <- as.data.frame(t(dplyr::select(counts_wide, -Gene)))

# paste together metadata and counts
meta_counts <- cbind(metaData, counts_t)

# only keep AP13, DAC, and F1's 
A_D <- filter(meta_counts, genotype == "AP13" | genotype == "DAC6" | genotype == "AxD_F1")

# change rownames to sample names 
rownames(A_D) <- A_D$Sample

# cleanup
AD_clean <- dplyr::select(A_D, -Library, -habitat, -genotype, -SRA.run, -SRA.study, -Sample)

# save colData for later
colData <- dplyr::select(A_D, habitat, genotype)

# transpose for next steps
AD_t <- as.data.frame(t(AD_clean))

# calc median
AD_t$median = apply(AD_t, 1, median, na.rm=T)

# only keep genes w/median expression > 3
AD_filt <- dplyr::select(filter(AD_t, median > 3), -median)

# write the data.
write.csv(AD_filt, "~/Dropbox/LowryLab/data/filtered_AP13_DAC6_RNA.csv")
write.csv(colData, "~/Dropbox/LowryLab/data/filtered_AP13_DAC6_RNA_metadata.csv")


```

# read in the processed data and filter
```{r AP13 and DAC6 only}

# only keeping AP13 and DAC in MI and TX; removing DAC TX 3 bc it's a contaminant

AD_filt <- read.csv("~/Dropbox/LowryLab/data/expression/filtered_AP13_DAC6_RNA.csv")
colData <- read.csv("~/Dropbox/LowryLab/data/expression/filtered_AP13_DAC6_RNA_metadata.csv")

rownames(AD_filt) <- AD_filt$X

# reformatting counts
AP_DAC <- dplyr::select(AD_filt, 
              PKLE_DAC_1,
              PKLE_AP13_1,
              PKLE_AP13_2,
              KBSM_AP13_1,
              KBSM_AP13_2,
              KBSM_DAC_1,
              KBSM_AP13_3,
              KBSM_AP13_4,
              KBSM_AP13_5,
              PKLE_DAC_4,
              KBSM_DAC_2,
              PKLE_AP13_3,
              PKLE_AP13_4,
              KBSM_DAC_3,
              PKLE_AP13_5)

# reformatting col Data
colData_AP_DAC <- filter(colData, 
       habitat == "KBSM" |
      habitat == "PKLE", 
      genotype == "DAC6" |
      genotype == "AP13") %>% slice(-9)

#write.csv(modules_all, "~/Dropbox/LowryLab/data/geneNames_modules.csv")
```

# run DESeq (make dds object)
```{r make a DESeq object}
# format for DESeq 

dds <- DESeqDataSetFromMatrix(countData = AP_DAC,
                              colData = colData_AP_DAC,
                              design =  ~ genotype + habitat + genotype:habitat)

```

# filter the DDS object
```{r filtering}
smallestGroupSize <- 2
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]

```

# normalize & run DESeq
```{r normalize & run}
# running deseq 
dds <- DESeq(dds)

# running VSD
vsd <- varianceStabilizingTransformation(dds)

```

# explore DESeq results to identify genes with habitat and genotype effects but no GxE
```{r explore DE results}

normCounts$gene <- rownames(normCounts)

expression_sig <- filter(normCounts, gene %in% rownames(GxEgenes))

# get the mean expression 
# PKLE
expression_sig <- expression_sig %>% 
  mutate(PKLE_DAC_mean = rowMeans(pick(PKLE_DAC_1, PKLE_DAC_4), na.rm = TRUE))

expression_sig <- expression_sig %>% 
  mutate(PKLE_AP13_mean = rowMeans(pick(PKLE_AP13_5, PKLE_AP13_4, PKLE_AP13_3), na.rm = TRUE))

# KBSM
expression_sig <- expression_sig %>% 
  mutate(KBSM_DAC_mean = rowMeans(pick(KBSM_DAC_1, KBSM_DAC_2, KBSM_DAC_3), na.rm = TRUE))

expression_sig <- expression_sig %>% 
  mutate(KBSM_AP13_mean = rowMeans(pick(KBSM_AP13_1, KBSM_AP13_2, KBSM_AP13_3, KBSM_AP13_4, KBSM_AP13_5), na.rm = TRUE))

# join the results stat to the expression means file
expMeans <- dplyr::select(expression_sig, gene, PKLE_DAC_mean, PKLE_AP13_mean, KBSM_DAC_mean, KBSM_AP13_mean)

expMeans$lFC <- filter(results_full, padj < 0.05)$log2FoldChange
expMeans$stat <- filter(results_full, padj < 0.05)$stat


# now scan for rank changing vs. variance changing
expMeans <- mutate(expMeans, PKLE_habDiff = PKLE_DAC_mean - KBSM_DAC_mean)
expMeans <- mutate(expMeans, PKLE_habDiff = PKLE_AP13_mean - KBSM_AP13_mean)

expMeans$sign <- expMeans$DAC_habDiff == expMeans$AP13_habDiff

ggplot(expMeans,
       aes(x = DAC_habDiff,
           y = AP13_habDiff))

ggplot(expMeans,
       aes(x = PKLE_DAC_mean,
           y = KBSM_DAC_mean))

write.csv(expMeans,
          "~/Dropbox/mean_expression.csv")

hab_geno <- (intersect(rownames(filter(results_genotype, padj < 0.05)), rownames(filter(results_habitat, padj < 0.05))))


site_geno_noGxE <- setdiff(hab_geno, rownames(filter(results_full, padj < 0.05)))

# now make a table to illustrate the 5000 genes with G and E, but no GxE 
site_gene_table <- filter(results_habitat,
       rownames(results_habitat) %in% site_geno_noGxE)
site_gene_table$model <- "expression ~ site"

geno_gene_table <- filter(results_genotype,
       rownames(results_genotype) %in% site_geno_noGxE)
geno_gene_table$model <- "expression ~ genotype"

site_geno_noGxE <- rbind(site_gene_table,
      geno_gene_table)


# write to save for supplement
write.csv(site_gene_table, "~/Dropbox/LowryLab/TableS2.csv")
```

# PCA plot of gene expression to vet samples
```{r PCA of genotypes}

pcaData <- plotPCA(vsd, intgroup=c("habitat", "genotype"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=habitat, shape=genotype)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme_bw() 


```

# export DE gene data 
```{r export }
# create and edit habitat contrast df
results_habitat <- as.data.frame(results(dds, name="habitat_PKLE_vs_KBSM"))
results_habitat$diffexpressed <- "NO"
results_habitat$diffexpressed[results_habitat$log2FoldChange > 1 & results_habitat$padj < 0.05] <- "UP"
results_habitat$diffexpressed[results_habitat$log2FoldChange < -1 & results_habitat$padj < 0.05] <- "DOWN"

# create and edit genotype contrast df
results_genotype <- as.data.frame(results(dds, name="genotype_DAC6_vs_AP13"))
results_genotype$diffexpressed <- "NO"
results_genotype$diffexpressed[results_genotype$log2FoldChange > 1 & results_genotype$padj < 0.05] <- "UP"
results_genotype$diffexpressed[results_genotype$log2FoldChange < -1 & results_genotype$padj < 0.05] <- "DOWN"

# create and edit full model contrast df
results_full <- as.data.frame(results(dds))
results_full$diffexpressed <- "NO"
results_full$diffexpressed[results_full$log2FoldChange > 1 & results_full$padj < 0.05] <- "UP"
results_full$diffexpressed[results_full$log2FoldChange < -1 & results_full$padj < 0.05] <- "DOWN"

write.csv(results_habitat, "~/Dropbox/LowryLab/data/expression/KBSM_PKLE_DE.csv")
write.csv(results_genotype, "~/Dropbox/LowryLab/data/expression/AP13_DAC6_DE.csv")
write.csv(results_full, "~/Dropbox/LowryLab/data/expression/full_DE.csv")

# get number of DE genes 
filter(results_full, padj < 0.05, diffexpressed == "DOWN" | diffexpressed == "UP")

# add gene name 
results_full$geneName <- rownames(results_full)

# separate gene name intu gene & chrom
results_labeled <- results_full %>%
  separate_wider_delim(
    geneName,
    delim = "G",
    names = c("chrom", "gene"),
    too_few = "debug"
  )


# K_subgenome
K_results <- filter(results_labeled,
                    chrom == "Pavir.1K" |
                    chrom == "Pavir.2K" |
                    chrom == "Pavir.3K" |
                      chrom == "Pavir.4K" |
                      chrom == "Pavir.5K" |
                      chrom == "Pavir.6K" |
                      chrom == "Pavir.7K" |
                      chrom == "Pavir.8K" |
                      chrom == "Pavir.9K")

# N_subgenome
N_results <- filter(results_labeled,
                    chrom == "Pavir.1N" |
                    chrom == "Pavir.2N" |
                    chrom == "Pavir.3N" |
                      chrom == "Pavir.4N" |
                      chrom == "Pavir.5N" |
                      chrom == "Pavir.6N" |
                      chrom == "Pavir.7N" |
                      chrom == "Pavir.8N" |
                      chrom == "Pavir.9N")

# add subgenome column
N_results$subgenome <- "N"
K_results$subgenome <- "K"

subGen_labeled <- rbind(N_results,
      K_results)

filter(subGen_labeled, padj < 0.05, abs(log2FoldChange) > 1, subgenome == "N")
filter(subGen_labeled, padj < 0.05, abs(log2FoldChange) > 1, subgenome == "K")

filter(subGen_labeled, padj < 0.05, abs(log2FoldChange) > 1, diffexpressed == "DOWN")


1593 + 1569

1569 / 3162
1593 / 3162

results_habitat
```

# make volcano plot of DE genes
```{r volcano plot}

# this can be done multiple times, for each of the models 
ggplot(annotDE, aes(x = log2FoldChange, y = -log10(padj), 
                    col = diffexpressed,
                    alpha = acid.phosphatases)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("NO" = "grey", "UP" = "red", "DOWN" = "blue")) +
  labs(title = "Full Model DE", x = "Log2 Fold Change", y = "-Log10 Adjusted P-value") +
  geom_vline(xintercept = c(-1, 1), col = "grey") + # Add vertical lines for LFC threshold
  geom_hline(yintercept = -log10(0.05), col = "grey") # Add horizontal line for p-value threshold

```

# upload necessary data for GO analyses 
```{r upload GO terms}
# read in the genes with GO terms from the annotation 
Pvirg_GOterms <- read.csv("~/Dropbox/LowryLab/data/expression/Pavir_GOterms_multiple.csv")

# melt to make longer
longer_GOterms <- unique(reshape2::melt(Pvirg_GOterms,
     id.vars=c("locusName")))

# convert to clusterProfiler format
GOterms_mult  <- as.data.frame(cbind(longer_GOterms$value,
                    longer_GOterms$locusName))

# extract names of GO terms
# Extract a named vector of all GO terms and their names
goterms <- Term(GOTERM)

# Convert to a data frame for easier use
term2name <- data.frame("GOID" = names(goterms), "term" = goterms)

```

```{r enrichment of other GO terms}
# enrichment of genotype + habitat genes (not GxE)
DE_site_geno <- enricher(rownames(site_geno_noGxE), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)



MidwestGxEGO <- enricher(as.vector(filter(Midwest_GxE_eQTL_maf, folded > 0.05)$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

GulfGxEGO <- enricher(as.vector(filter(Gulf_GxE_eQTL_maf, folded > 0.05)$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

AtlanticGxEGO <- enricher(as.vector(filter(Atlantic_GxE_eQTL_maf, folded > 0.05)$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

dotplot(DE_site_geno) + ggtitle("GO enrichment, site + genotype DE genes")
```

# format DE genes for GO term analyses
```{r format for GO}

results_full$gene <- gsub(".v6.1", "", rownames(results_full))
results_genotype$gene <- gsub(".v6.1", "", rownames(results_genotype))
results_habitat$gene <- gsub(".v6.1", "", rownames(results_habitat))


```

# go analysis of DE genes 
```{r run GO analyses of DE genes}

results_full <- read.csv("~/Dropbox/LowryLab/data/expression/full_DE.csv")

# full model DE
fullMod <- enricher(as.vector(filter(results_full, diffexpressed!="NO")$X), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# habitat DE
habMod <- enricher(as.vector(filter(results_habitat, diffexpressed!="NO")$gene), 
                   universe = Pvirg_GOterms,
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# genotype DE 
genMod <- enricher(as.vector(filter(results_genotype, diffexpressed!="NO")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# write the results
write.csv(as.data.frame(fullMod@result), "~/Dropbox/LowryLab/data/expression/GO_full_DE.csv")
write.csv(as.data.frame(habMod@result), "~/Dropbox/LowryLab/data/expression/GO_habitat_DE.csv")
write.csv(as.data.frame(genMod@result), "~/Dropbox/LowryLab/data/expression/GO_genotype_DE.csv")

```

# prep for WGCNA
```{r WGCNA prep}

# extract the normalized data for WGCNA and format 
wpn_vsd <- getVarianceStabilizedData(dds)
rv_wpn <- rowVars(wpn_vsd)

#q75_wpn <- quantile( rowVars(wpn_vsd), .60) 
##expr_normalized <- wpn_vsd[ rv_wpn > q75_wpn, ]
#dim(expr_normalized)

# transpose
input_mat_all = t(wpn_vsd)

#write.csv(wpn_vsd, "~/Dropbox/LowryLab/data/expression/AP13_DAC6_vsdNorm.csv")


```

# pick the power level & run WGCNA for all genes; choose top hub genes  
```{r WGCNA all genes pick power and run}
# skipping straight to reading in 
wpn_vsd <- read.csv("~/Dropbox/LowryLab/data/expression/AP13_DAC6_vsdNorm.csv")
rownames(wpn_vsd) <- wpn_vsd$X
input_mat_all = t(dplyr::select(wpn_vsd, -X))


#> Allowing multi-threading with up to 4 threads
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 20, by = 2))

# Call the network topology analysis function -- now calling with the ~10000 genes with the most variance
sft = pickSoftThreshold(
  input_mat_all,             # <= Input data
  blockSize = 1000,
  powerVector = powers,
  verbose = 5
  )


# plot to select threshold
par(mfrow = c(1,2));
cex1 = 0.9;

plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)

#picked_power = 5
temp_cor <- cor
cor <- WGCNA::cor

```

# now run the network calls for WGCNA
```{r netwk call}

# generate the network
netwk_all <- blockwiseModules(input_mat_all,       
                          power = 7,           
                          networkType = "signed",
                          deepSplit = 2,
                          pamRespectsDendro = F,
                          minModuleSize = 100,
                          maxBlockSize = 4000,
                          reassignThreshold = 0,
                          mergeCutHeight = 0.3,
                          saveTOMs = T,
                          saveTOMFileBase = "ER",
                          numericLabels = F,
                          verbose = 3)




# print the network so you get the number of modules
#summary(as.factor(netwk_all$colors))

# calc the number of genes in each module with the cut height of 0.3

# format the data frame
#names_cols <- as.data.frame(netwk_all$colors)
#names_cols$gene <- rownames(names_cols)

# plot
#ggplot(names_cols,
#       aes(x = `netwk_all$colors`)) + 
#  geom_bar()


# choose top hub gene
#chooseTopHubInEachModule(
#   input_mat_all, 
#   netwk_all$colors, 
#   power = picked_power)

write.csv(as.data.frame(netwk_all$colors), "~/Desktop/bergetal_2026/WGCNA_moduleList.csv")

write.csv(netwk_all$MEs, "~/Desktop/bergetal_2026/module_eigengenes.csv")
```

# plot WGCNA for all genes
```{r plot WGCNA all }
 # Convert labels to colors for plotting


mergedColors_all = labels2colors(netwk_all$colors)

# using a higher cut height to merge into a more manageable number of modules
# this is old so I'm commenting it out
# mergeNet <- mergeCloseModules(input_mat_all, mergedColors_all, cutHeight=0.3)
# mergeCols <- labels2colors(mergeNet$colors)

# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  netwk_all$dendrograms[[1]],
  mergedColors_all[netwk_all$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )

module_df <- data.frame(
  gene_id = names(netwk_all$colors),
  colors = labels2colors(netwk_all$colors)
)

 
# Get Module Eigengenes per cluster
MEs0_all <- moduleEigengenes(input_mat_all, netwk_all$colors)$eigengenes

# Reorder modules so similar modules are next to each other
MEs0_all <- orderMEs(MEs0_all)
module_order_all = names(MEs0_all) %>% gsub("ME","", .)

# Add treatment names
MEs0_all$treatment = row.names(MEs0_all)

# tidy data
mME_all = MEs0_all %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order_all)
  )

# plot data 
mME_all %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")

dim(input_mat_all)
netwk_all

```

# get kME of all the modules (this is how you get the hub genes)
```{r kME hub genes}
hubDat <- signedKME(input_mat_all, 
  moduleEigengenes(input_mat_all, netwk_all$colors)$eigengenes, 
  exprWeights = NULL,
  MEWeights = NULL,
  outputColumnName = "kME", 
  corFnc = "cor", 
  corOptions = "use = 'p'")

write.csv(hubDat, "~/Dropbox/LowryLab/data/hubGenes.csv")
```

# formatting WGCNA output for GO analyses
```{r formatting module output}

modules_all <- as.data.frame(netwk_all$colors)
modules_all$gene <- gsub(".v6.1", "", rownames(modules_all))
modules_all$color <- modules_all$`netwk_all$colors`

#write.csv(modules_all, "~/Dropbox/LowryLab/data/genes_modules.csv")

```

# GO analyses for the modules 
```{r run GO analyses of the modules}
modules_all <- read.csv("~/Dropbox/LowryLab/data/genes_modules.csv")

tanMod <- enricher(as.vector(filter(modules_all, color == "tan")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

magentaMod <- enricher(as.vector(filter(modules_all, color == "magenta")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


turquoiseMod <- enricher(as.vector(filter(modules_all, color == "turquoise")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


salmonMod <- enricher(as.vector(filter(modules_all, color == "salmon")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


purpleMod <- enricher(as.vector(filter(modules_all, color == "purple")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


grey60Mod <- enricher(as.vector(filter(modules_all, color == "grey60")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


greenMod <- enricher(as.vector(filter(modules_all, color == "green")$gene), 
                     GOterms_mult$V2,
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         #minGSSize = 10, 
         #maxGSSize = 500, 
         qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

yellowMod <- enricher(as.vector(filter(modules_all, color == "yellow")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

lightcyanMod <- enricher(as.vector(filter(modules_all, color == "lightcyan")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

pinkMod <- enricher(as.vector(filter(modules_all, color == "pink")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

greenyellowMod <- enricher(as.vector(filter(modules_all, color == "greenyellow")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

brownMod <- enricher(as.vector(filter(modules_all, color == "brown")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

blackMod <- enricher(as.vector(filter(modules_all, color == "black")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

redMod <- enricher(as.vector(filter(modules_all, color == "red")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


midnightblueMod <- enricher(as.vector(filter(modules_all, color == "midnightblue")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

cyanMod <- enricher(as.vector(filter(modules_all, color == "cyan")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


blueMod <- enricher(as.vector(filter(modules_all, color == "blue")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

lightyellowMod <- enricher(as.vector(filter(modules_all, color == "lightyellow")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


dotplot(tanMod) + ggtitle("Tan Module GO Term Enrichment")
dotplot(magentaMod) + ggtitle("Magenta Module GO Term Enrichment")
dotplot(turquoiseMod) + ggtitle("Turquoise Module GO Term Enrichment")
dotplot(salmonMod) + ggtitle("Salmon Module GO Term Enrichment")
dotplot(purpleMod) + ggtitle("Purple Module GO Term Enrichment")
dotplot(grey60Mod) + ggtitle("Grey60 Module GO Term Enrichment")
dotplot(greenMod) + ggtitle("Green Module GO Term Enrichment")
dotplot(yellowMod) + ggtitle("Yellow Module GO Term Enrichment")
#dotplot(lightcyanMod) + ggtitle("Light Cyan Module GO Term Enrichment")
#dotplot(pinkMod) + ggtitle("Pink Module GO Term Enrichment")
#dotplot(greenyellowMod) + ggtitle("GreenYellow Module GO Term Enrichment")
dotplot(brownMod) + ggtitle("Brown Module GO Term Enrichment")
dotplot(blackMod) + ggtitle("Black Module GO Term Enrichment")
dotplot(redMod) + ggtitle("Red Module GO Term Enrichment")
dotplot(midnightblueMod) + ggtitle("Midnight Blue Module GO Term Enrichment")
dotplot(cyanMod) + ggtitle("Cyan Blue Module GO Term Enrichment")
dotplot(blueMod) + ggtitle("Blue Module GO Term Enrichment")

dotplot(lightyellowMod) + ggtitle("Light Module GO Term Enrichment")
```

# GO analyses for modules (all genes)
```{r}
Mod_allGenes <- enricher(as.vector(filter(modules_all, color == "green")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE = test_mult, TERM2NAME = term2name)

write.csv(ME_labeled, "~/Dropbox/LowryLab/ME_labeled.csv")
```


```{r}
mME_all <- read.csv("~/Dropbox/LowryLab/ME_labeled.csv")
mME_all

mME_all %>% ggplot(., aes(x=treatment, y=X, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")

```



```{r module-trait correlations}
# formatting the Mes0 datasheet to examine module - trait correlations 
MEs0_all$sample <- rownames(MEs0_all)

PKLEs <- filter(MEs0_all, sample == "PKLE_DAC_1" |
                  sample == "PKLE_AP13_1" | 
                  sample == "PKLE_AP13_2" | 
                  sample == "PKLE_DAC_4" |
                  sample ==  "PKLE_AP13_5")

KBSMs <- filter(MEs0_all, sample == "KBSM_AP13_1"|
                   sample ==  "KBSM_AP13_2" |
                   sample == "KBSM_DAC_1" |
                   sample ==  "KBSM_AP13_3" |
                   sample == "KBSM_AP13_4"|
                   sample == "KBSM_AP13_5"|
                   sample ==  "KBSM_DAC_2" |
                  sample ==   "KBSM_DAC_3" )

PKLEs$habitat <- "PKLE"
KBSMs$habitat <- "KBSM"

habLab <- rbind(PKLEs,
      KBSMs)

APs <- filter(habLab, sample == "KBSM_AP13_1"|
                   sample ==  "KBSM_AP13_2" |
                   sample ==  "KBSM_AP13_3" |
                   sample == "KBSM_AP13_4"|
                   sample == "KBSM_AP13_5"| 
                  sample == "PKLE_AP13_1" | 
                  sample == "PKLE_AP13_2" | 
                  sample ==  "PKLE_AP13_5")

DACs <- filter(habLab, sample == "KBSM_DAC_1" |
                   sample ==  "KBSM_DAC_2" |
                  sample ==  "KBSM_DAC_3" |
                  sample == "PKLE_DAC_4" |
                 sample == "PKLE_DAC_1")

APs$genotype <- "AP13"
DACs$genotype <- "DAC6"

ME_labeled <- rbind(APs,
                    DACs)


summary(aov(MEtan ~ habitat * genotype, ME_labeled))
summary(aov(MEmagenta ~ habitat * genotype, ME_labeled))
summary(aov(MEturquoise ~ habitat * genotype, ME_labeled))
summary(aov(MEsalmon ~ habitat * genotype, ME_labeled))
summary(aov(MEpurple ~ habitat * genotype, ME_labeled))
summary(aov(MEgrey60 ~ habitat * genotype, ME_labeled))
summary(aov(MEgreen ~ habitat * genotype, ME_labeled))
summary(aov(MEyellow ~ habitat * genotype, ME_labeled))
summary(aov(MElightcyan ~ habitat * genotype, ME_labeled))
summary(aov(MEpink ~ habitat * genotype, ME_labeled))
summary(aov(MEgreenyellow ~ habitat * genotype, ME_labeled))
summary(aov(MEbrown~ habitat * genotype, ME_labeled))
summary(aov(MEblack~ habitat * genotype, ME_labeled))
summary(aov(MEred~ habitat * genotype, ME_labeled))
summary(aov(MEmidnightblue~ habitat * genotype, ME_labeled))
summary(aov(MEcyan~ habitat * genotype, ME_labeled))
summary(aov(MEblue~ habitat * genotype, ME_labeled))

summary(aov(MEtan ~ genotype, ME_labeled))
summary(aov(MEtan ~ habitat, ME_labeled))
summary(aov(MEtan ~ habitat * genotype, ME_labeled))

summary(aov(MEmagenta ~ genotype, ME_labeled))
summary(aov(MEmagenta ~ habitat, ME_labeled))

summary(aov(MEturquoise ~ genotype, ME_labeled))
summary(aov(MEturquoise ~ habitat, ME_labeled))

summary(aov(MElightgreen ~ genotype, ME_labeled))
summary(aov(MElightgreen ~ habitat, ME_labeled))

summary(aov(MEsalmon ~ genotype, ME_labeled))
summary(aov(MEsalmon ~ habitat, ME_labeled))

summary(aov(MEpurple ~ genotype, ME_labeled))
summary(aov(MEpurple ~ habitat, ME_labeled))

summary(aov(MEgrey60 ~ genotype, ME_labeled))
summary(aov(MEgrey60 ~ habitat, ME_labeled))

summary(aov(MEgreen ~ genotype, ME_labeled))
summary(aov(MEgreen ~ habitat, ME_labeled))

summary(aov(MEyellow ~ genotype, ME_labeled))
summary(aov(MEyellow ~ habitat * genotype, ME_labeled))

summary(aov(MElightcyan ~ genotype, ME_labeled))
summary(aov(MElightcyan~ habitat, ME_labeled))

summary(aov(MEpink ~ genotype, ME_labeled))
summary(aov(MEpink ~ habitat, ME_labeled))

summary(aov(MElightyellow ~ genotype, ME_labeled))
summary(aov(MElightyellow ~ habitat, ME_labeled))

summary(aov(MEgreenyellow ~ genotype, ME_labeled))
summary(aov(MEgreenyellow ~ habitat, ME_labeled))

summary(aov(MEbrown ~ genotype, ME_labeled))
summary(aov(MEbrown ~ habitat, ME_labeled))
summary(aov(MEbrown ~ habitat * genotype, ME_labeled))

summary(aov(MEblack ~ genotype, ME_labeled))
summary(aov(MEblack ~ habitat, ME_labeled))

summary(aov(MEred ~ genotype, ME_labeled))
summary(aov(MEred ~ habitat, ME_labeled))

summary(aov(MEmidnightblue ~ genotype, ME_labeled))
summary(aov(MEmidnightblue ~ habitat, ME_labeled))

summary(aov(MEcyan ~ genotype, ME_labeled))
summary(aov(MEcyan ~ habitat, ME_labeled))

```


# Differential network analysis between genotypes -----------------------

```{r}

# Optional: Perform differential network analysis between genotypes
# Split the data by genotype

AP13 <- dplyr::select(as.data.frame(expr_normalized), PKLE_AP13_1, PKLE_AP13_2, PKLE_AP13_3, KBSM_AP13_1, KBSM_AP13_2, KBSM_AP13_3, KBSM_AP13_4, KBSM_AP13_5)
DAC <- dplyr::select(as.data.frame(expr_normalized), KBSM_DAC_1, KBSM_DAC_2, KBSM_DAC_3, PKLE_DAC_1, PKLE_DAC_4)


#[1] "PKLE_DAC_1" "KBSM_AP13_1" "KBSM_AP13_2" "KBSM_DAC_1"  "KBSM_AP13_3" "KBSM_AP13_4" "KBSM_AP13_5" "PKLE_DAC_4"  "KBSM_DAC_2" 
#[13] "PKLE_AP13_4" "KBSM_DAC_3"  "PKLE_AP13_5"

# Run module preservation analysis
# This checks if modules identified in one genotype are preserved in the other

#multiExpr <- list(Coast = list(data = datExpr_coast), Inland = list(data = datExpr_inland))

gsg_AP13 = goodSamplesGenes(t(AP13), verbose = 3)
gsg_DAC6 = goodSamplesGenes(t(DAC), verbose = 3)

AP13_filt = t(AP13)[gsg_AP13$goodSamples, gsg_AP13$goodGenes]
DAC6_filt = t(DAC)[gsg_DAC6$goodSamples, gsg_DAC6$goodGenes]

multiExpr <- list(AP13 = list(data = AP13_filt), DAC6 = list(data = DAC6_filt))
multiColot <- list(AP13 = list(data = netwk_all$colors), DAC6 = list(data = netwk_all$colors))
multiColot
# Calculate module preservation statistics
mp <- modulePreservation(multiExpr, 
                         networkType = "signed", 
                         referenceNetworks = c(1), 
                         nPermutations = 100, # Increase for publication-quality analysis
                         randomSeed = 12345, 
                         verbose = 3,
                         mergedColors_all,
                         netwk_all$colors,
                         multiWeights = NULL,
                         dataIsExpr = TRUE)

multiExpr_AP13

# Create a table with the preservation statistics
ref <- 1 # Reference network (Coast)
test <- 2 # Test network (Inland)

# Extract preservation statistics
# Z-summary is a measure of preservation
preservation_stats <- data.frame(
  Module = modNames,
  moduleSize = mp$quality$moduleSize[[ref]],
  Zsummary = mp$preservation$Z$ref.vs.test[[ref]][[test]]$Zsummary$observed,
  medianRank = mp$preservation$rankPres$ref.vs.test[[ref]][[test]]$medianRank$observed
)

# Sort the modules by Z-summary
preservation_stats <- preservation_stats[order(preservation_stats$Zsummary, decreasing = TRUE), ]

# Write the preservation statistics to a file
write.csv(preservation_stats, file = "module_preservation_statistics.csv", row.names = FALSE)

# Plot the preservation statistics
pdf("module_preservation_plot.pdf", width = 10, height = 8)
par(mfrow = c(1, 2))
# Z-summary plot
x <- mp$preservation$Z$ref.vs.test[[ref]][[test]]$Zsummary$observed
y <- mp$preservation$Z$ref.vs.test[[ref]][[test]]$moduleSize
plot(y, x, col = moduleColors, 
     main = "Module preservation - Zsummary",
     xlab = "Module size", ylab = "Zsummary",
     pch = 19, cex = 1.5)
abline(h = 2, col = "blue", lty = 2)
abline(h = 10, col = "red", lty = 2)
text(y, x, labels = modNames, pos = 1, cex = 0.7)
# Median rank plot
y <- mp$preservation$rankPres$ref.vs.test[[ref]][[test]]$medianRank$observed
plot(mp$preservation$rankPres$ref.vs.test[[ref]][[test]]$moduleSize, y,
     col = moduleColors, 
     main = "Module preservation - Median rank",
     xlab = "Module size", ylab = "Median rank",
     pch = 19, cex = 1.5)
text(mp$preservation$rankPres$ref.vs.test[[ref]][[test]]$moduleSize, y,
     labels = modNames, pos = 1, cex = 0.7)
dev.off()

```

