---
title: "DE_analyses and WGCNA"
author: "Colette Berg"
date: "2025-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(edgeR)
library(DESeq2)
library(genefilter)
library(tidyverse)
library(WGCNA)
library(gridExtra)

setwd("~/Dropbox/LowryLab/data/DE_analyses")
```

# read in the data 
```{r read in data}
# read in read counts from HTSeq; 
# not including the F1's 

# BUWCG <-  read.csv("/Users/coletteberg/Dropbox/LowryLab/data/BUWCG.counts.tsv")
BUWCU <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWCU.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWGW <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWGW.counts.tsv", sep="\t", header=FALSE, row.names = 1)
#BUWGX <-  read.csv("BUWGX.counts.tsv")
BUWHN <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWHN.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWBN <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWBN.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWBY <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWBY.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWNS <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWNS.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWOC <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWOC.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWOP <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWOP.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWGG <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWGG.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWNH <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWNH.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWOO <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWOO.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWBH <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWBH.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWGY <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWGY.counts.tsv", sep="\t", header=FALSE, row.names = 1)
BUWHP <-  read.csv("~/Dropbox/LowryLab/data/DE_analyses/BUWHP.counts.tsv", sep="\t", header=FALSE, row.names = 1)
#BUWBG <-  read.csv("BUWBG.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWBU <-  read.csv("BUWBU.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWBW <-  read.csv("BUWBW.counts.tsv")
#BUWCY <-  read.csv("BUWCY.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWGO <-  read.csv("BUWGO.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWHS <-  read.csv("BUWHS.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWHW <-  read.csv("BUWHW.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWHY <-  read.csv("BUWHY.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWHZ <-  read.csv("BUWHZ.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWNA <-  read.csv("BUWNA.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
#BUWNP <-  read.csv("BUWNP.counts.tsv", sep="\t", header=FALSE, , row.names = 1)
# BUWNX <-  read.csv("BUWNX.counts.tsv")

# put all of the files together
all <- cbind(BUWCU,
BUWGW,
BUWHN,
BUWBN,
BUWBY,
BUWNS,
BUWOC,
BUWOP,
BUWGG,
BUWNH,
BUWOO,
BUWBH,
#BUWGY,
BUWHP)


# set the colnames
colnames(all) <- c("AP13_KBSM_1",
"AP13_KBSM_2",
"AP13_KBSM_3",
"AP13_PKLE_1",
"AP13_PKLE_2",
"AP13_PKLE_3",
"AP13_PKLE_4",
"AP13_PKLE_5",
"DAC6_KBSM_1",
"DAC6_KBSM_2",
"DAC6_KBSM_3",
"DAC6_PKLE_1",
#"DAC6_PKLE_2",
"DAC6_PKLE_3")


# read sample data 
colData <- read.csv("~/Dropbox/LowryLab/data/DE_analyses/sampleDat_noF1_1.csv",
                    row.names = 1)

```

# run DESeq (make dds object)
```{r make a DESeq object}
# format for DESeq 
# the first model includes genotype, habitat, and their interaction 
dds <- DESeqDataSetFromMatrix(countData = all,
                              colData = colData,
                              design =  ~ habitat + genotype + habitat:genotype)

```

# filter the DDS object
```{r filter}



# filter for each (all model, hab only, gen only)

keep <- rowSums(counts(dds) >= 10) >= 2
dds <- dds[keep,]

```


# normalize
```{r normalize }

# running deseq for habitat only
dds <- DESeq(dds)
vsd <- varianceStabilizingTransformation(dds)

```

```{r}
res_genotype <- results(dds, contrast=c("genotype","AP13","DAC6"))

res_habitat <- results(dds, contrast=c("habitat","KBSM","PKLE"))

```


# PCA plot of gene expression to vet samples

```{r PCA of genotypes}

pcaData <- plotPCA(vsd, intgroup=c("habitat", "genotype"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=habitat, shape=genotype)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme_bw()
```
# plotting differential gene expression 

```{r volcano plots}


# DE based on genotype
gen_DE <- ggplot(res_genotype,
       aes(x=log2FoldChange,
           y=-log10(padj))) +
  geom_point(size = 0.35) +
  theme_bw() +
  ggtitle("AP13 vs DAC6 DE")

# DE based on habitat
hab_DE <- ggplot(res_habitat,
       aes(x=log2FoldChange,
           y=-log10(padj))) +
  geom_point(size = 0.35) +
  theme_bw() +
  theme(legend.position = "none") +
  ggtitle("MI vs TX DE")


# putting the two plots together
grid.arrange(gen_DE,
             hab_DE,
             nrow=1)

```


```{r prep for WGCNA}
# just doing WGCNA on the whole model (genotype, habitat, their interaction)

# extract the normalized data 
wpn_vsd <- getVarianceStabilizedData(dds)
rv_wpn <- rowVars(wpn_vsd)

# Extracting the 5% most differentiated genes
# q95_wpn <- quantile( rowVars(wpn_vsd), .05) 

# set the row variables
# expr_normalized <- wpn_vsd[ rv_wpn > q95_wpn, ]

# format for WGCNA
input_mat = t(wpn_vsd)

```


# pick the power level for WGCNA

```{r WGCNA picking power}

allowWGCNAThreads()

#> Allowing multi-threading with up to 4 threads
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 20, by = 2))

# Call the network topology analysis function
sft = pickSoftThreshold(
  input_mat,             # <= Input data
  #blockSize = 30,
  powerVector = powers,
  verbose = 5
  )


# plot to select threshold
par(mfrow = c(1,2));
cex1 = 0.9;

plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)
     
```

# run the heavy WGCNA analysis
```{r WGCNA analysis }
picked_power = 12
temp_cor <- cor
cor <- WGCNA::cor


netwk <- blockwiseModules(input_mat,                # <= input here

                          # == Adjacency Function ==
                          power = picked_power,                # <= power here
                          networkType = "signed",

                          # == Tree and Block Options ==
                          deepSplit = 2,
                          pamRespectsDendro = F,
                          # detectCutHeight = 0.75,
                          minModuleSize = 50,
                          maxBlockSize = 4000,

                          # == Module Adjustments ==
                          reassignThreshold = 0,
                          mergeCutHeight = 0.25,

                          # == TOM == Archive the run results in TOM file (saves time)
                          saveTOMs = T,
                          saveTOMFileBase = "ER",

                          # == Output Options
                          numericLabels = T,
                          verbose = 3)


```

# dynamic tree cutting
```{r dynamic tree cutting}

softPower = 12 #Chosen in the graphs before
adjacency = adjacency(input_mat, power = softPower, type = "unsigned") #Calculating the adjacency matrix
#help(adjacency )

mem.maxVSize()
```



# plot the WGCNA dendrogram

```{r WGCNA plotting dendrogram}
 # Convert labels to colors for plotting
mergedColors = labels2colors(netwk$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )

MEList <- moduleEigengenes(input_mat, colors = mergedColors)
MEs <- MEList$eigengenes

# Calculate dissimilarity of module eigengenes
MEDiss <- 1 - cor(MEs)
# Cluster module eigengenes
METree <- hclust(as.dist(MEDiss), method = "average")

MEDissThres <- 0.25

# merging similar modules 
merge <- mergeCloseModules(input_mat, mergedColors, cutHeight = MEDissThres, verbose = 3)


# The merged module colors
mergedColors <- merge$colors

# Eigengenes of the new merged modules
mergedMEs <- merge$newMEs

MEs0 <- orderMEs(MEs0)
module_order = names(mergedMEs) %>% gsub("ME","", .)

# Add treatment names
mergedMEs$treatment = row.names(mergedMEs)

# tidy & plot data
mME = mergedMEs %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )

mME %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")


mME = mergedMEs %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )

# make 
```

# plot the WGCNA module heat map 

```{r WGCNA plotting heat map of modules}
module_df <- data.frame(
  gene_id = names(netwk$colors),
  colors = labels2colors(netwk$colors)
)


write_delim(module_df,
            file = "gene_modules.txt",
            delim = "\t")

# Get Module Eigengenes per cluster
MEs0 <- moduleEigengenes(input_mat, mergedColors)$eigengenes

# Reorder modules so similar modules are next to each other
MEs0 <- orderMEs(MEs0)
module_order = names(MEs0) %>% gsub("ME","", .)

# Add treatment names
MEs0$treatment = row.names(MEs0)

# tidy & plot data
mME = MEs0 %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )

mME %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")
```

```{r get module expression value}
# pick out a few modules of interest here
modules_of_interest = c("turquoise", "yellow")

# Pull out list of genes in that module
submod = module_df %>%
  subset(colors %in% modules_of_interest)

row.names(module_df) = module_df$gene_id

# Get normalized expression for those genes

subexpr = wpn_vsd[submod$gene_id,]

submod_df = data.frame(subexpr) %>%
  mutate(
    gene_id = row.names(.)
  ) %>%
  pivot_longer(-gene_id) %>%
  mutate(
    module = module_df[gene_id,]$colors
  )

submod_df

ggplot(filter(submod_df),
              aes(x=name, y=value)) +
  geom_boxplot() +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90)
  ) +
  facet_grid(rows = vars(module)) +
  labs(x = "treatment",
       y = "normalized expression") +
  facet_wrap(~module) + ylim(3,12)

write.csv(submod_df, "submod_geneExp.csv")

dat <- read.csv("submod_geneExp.csv")

ggplot(filter(dat, module == "turquoise" | module == "yellow"),
       aes(x = name,
           y = value,
           col = hab)) + 
  geom_boxplot() +
  facet_wrap(~module) +
  ylim(3.5,12) + 
  theme_bw() +
  scale_color_manual(values = c("darkorange",
                                "darkgreen"))

```



# identify the genes with the highest hub membership (from David's code)

```{r hub gene identification}

```

# ASE time 
```{r ASE time}
# read in the outputs from GATK
KBSM_AxD_1 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWBG.table", sep = "\t")
KBSM_AxD_2 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWBU.table", sep = "\t")
#KBSM_AxD_3 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWBW.table", sep = "\t")
KBSM_AxD_4 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWHW.table", sep = "\t")
KBSM_AxD_5 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWHY.table", sep = "\t")
KBSM_AxD_6 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWNX.table", sep = "\t")
PKLE_AxD_1 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWCY.table", sep = "\t")
PKLE_AxD_2 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWGO.table", sep = "\t")
PKLE_AxD_3 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWHS.table", sep = "\t")
PKLE_AxD_4 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWHZ.table", sep = "\t")
PKLE_AxD_5 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWNA.table", sep = "\t")
PKLE_AxD_6 <- read.csv("~/Dropbox/LowryLab/data/ASE/BUWNP.table", sep = "\t")

# filter to only include sites with sufficient depth to test for ASE 
# KBSM
KBSM_AxD_1_filt <- filter(KBSM_AxD_1, totalCount > 30, totalCount < 300)
KBSM_AxD_2_filt <- filter(KBSM_AxD_2, totalCount > 30, totalCount < 300)
KBSM_AxD_4_filt <- filter(KBSM_AxD_4, totalCount > 30, totalCount < 300)
KBSM_AxD_5_filt <- filter(KBSM_AxD_5, totalCount > 30, totalCount < 300)
KBSM_AxD_6_filt <- filter(KBSM_AxD_6, totalCount > 30, totalCount < 300)

# PKLE
PKLE_AxD_1_filt <- filter(PKLE_AxD_1, totalCount > 30, totalCount < 300)
PKLE_AxD_2_filt <- filter(PKLE_AxD_2, totalCount > 30, totalCount < 300)
PKLE_AxD_3_filt <- filter(PKLE_AxD_3, totalCount > 30, totalCount < 300)
PKLE_AxD_4_filt <- filter(PKLE_AxD_4, totalCount> 30, totalCount < 300)
PKLE_AxD_5_filt <- filter(PKLE_AxD_5, totalCount > 30, totalCount < 300)
PKLE_AxD_6_filt <- filter(PKLE_AxD_6, totalCount > 30, totalCount < 300)


```

```{r binomial test }

# MI sample 1
KBSM_F1_1_tested <- KBSM_AxD_1_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# MI sample 2
KBSM_F1_2_tested <- KBSM_AxD_2_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# MI sample 3
KBSM_F1_4_tested <- KBSM_AxD_4_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# MI sample 4 
KBSM_F1_5_tested <- KBSM_AxD_5_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# MI sample 5 
KBSM_F1_6_tested <- KBSM_AxD_6_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 


# TX sample 1 
PKLE_F1_1_tested <- PKLE_AxD_1_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# TX sample 2
PKLE_F1_2_tested <- PKLE_AxD_2_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# TX sample 3
PKLE_F1_3_tested <- PKLE_AxD_3_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# TX sample 4
PKLE_F1_4_tested <- PKLE_AxD_4_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# TX sample 5
PKLE_F1_5_tested <- PKLE_AxD_5_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

# TX sample 6
PKLE_F1_6_tested <- PKLE_AxD_6_filt %>%
  rowwise() %>%
  mutate(p_value = binom.test(refCount, totalCount, p = 0.5)$p.value) %>%
  ungroup() # 

class(KBSM_AxD_1_filt$totalCount)
class(KBSM_AxD_1_filt$refCount)

ggplot(KBSM_F1_2_tested,
       aes(x = position,
           y = -log10(p_value))) + 
  facet_wrap(~contig) + 
  geom_point()

filter(PKLE_F1_6_tested, p_value < 0.0001)

filter(PKLE_F1_5_tested, p_value < 0.0001)


# find sites that demonstrate ASE in all 5 replicates 
#Reduce(intersect(
  
intersect(list(dplyr::select(filter(PKLE_F1_6_tested, p_value < 0.0001), contig, position),
          dplyr::select(filter(PKLE_F1_5_tested, p_value < 0.0001), contig, position),
          dplyr::select(filter(PKLE_F1_4_tested, p_value < 0.0001), contig, position),
          dplyr::select(filter(PKLE_F1_3_tested, p_value < 0.0001), contig, position),
          dplyr::select(filter(PKLE_F1_2_tested, p_value < 0.0001), contig, position),
          dplyr::select(filter(PKLE_F1_1_tested, p_value < 0.0001), contig, position)))
         
```

```{r}
KBSM_F1_1_refBias <- filter(KBSM_F1_1_tested, 
       p_value < 0.0000000001,
       refCount > altCount)

KBSM_F1_1_altBias <- filter(KBSM_F1_1_tested, 
       p_value < 0.0000000001,
       refCount < altCount)

PKLE_F1_1_refBias 

PKLE_F1_1_altBias
KBSM_F1_1_altBias
KBSM_F1_1_refBias

KBS_alt <- ggplot(KBSM_F1_1_altBias,
       aes(x = position, 
           y = contig)) +
  geom_point()  + 
  ggtitle("ASE biased to alternate allele, KBSM AxD F1")

KBS_ref <- ggplot(KBSM_F1_1_refBias,
       aes(x = position, 
           y = contig)) +
  geom_point() + 
  ggtitle("ASE biased to reference allele, KBSM AxD F1")

grid.arrange(KBS_alt, KBS_ref, nrow = 1)


```

