---
title: "DE_analyses and WGCNA"
author: "Colette Berg"
date: "2025-10-07"
output: html_document
---

# set up 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(edgeR)
library(DESeq2)
library(genefilter)
library(tidyverse)
library(WGCNA)
library(clusterProfiler)
library(GO.db)
library(reshape2)
library(fs)

setwd("~/Dropbox/LowryLab/data")
```


# read in the data 
```{r read in data}
# read in read counts from HTSeq; 
# not including the F1's 

# get a list of all the gene counts files 
tsv_files <- dir_ls(path = "~/Dropbox/LowryLab/data/HTSeq_Counts_pseudo/", regexp = "\\.tsv$")

# read all of them into one file
all_data <- read_tsv(tsv_files, id = "file_source", col_names = FALSE) # The 'id' argument adds a column with the filename

# clean up the individual ID column
all_data$file_source <- gsub("/Users/coletteberg/Dropbox/LowryLab/data/HTSeq_Counts_pseudo/", "", all_data$file_source)
all_data$file_source <- gsub(".geneID.counts.tsv", "", all_data$file_source)

# clean up the gene ID column
all_data$X1 <- gsub(".v6.1", "", all_data$X1)

# now add a genome column by string-splitting
allData_sep <- separate_wider_delim(all_data, X1, delim="_", names=c("Genome", "Gene"), too_many = "debug")
allData_clean <- filter(allData_sep, X1_pieces == 2)

# split by genome
AP13_counts <- filter(dplyr::select(allData_clean, -X1_ok, -X1_pieces, -X1_remainder, -X1), Genome == "AP13")
DAC6_counts <- filter(dplyr::select(allData_clean, -X1_ok, -X1_pieces, -X1_remainder, -X1), Genome == "DAC6")

# clean up count column
AP13_counts  <- rename(AP13_counts, AP13_count = X2)
DAC6_counts  <- rename(DAC6_counts, DAC6_count = X2)

# paste together
both_counts <- cbind(AP13_counts, dplyr::select(DAC6_counts, DAC6_count))

# sum up the counts
both_counts <- mutate(both_counts, counts = AP13_count + DAC6_count)

# keep the necessary columns
counts_clean_long <- dplyr::select(both_counts, file_source, Gene, counts)

# clean up name 
counts_clean_long <- rename(counts_clean_long, indv = file_source)

# pivot wider 
counts_wide <- pivot_wider(counts_clean_long, names_from=indv, values_from=counts)

# read in metadata 
metaData <- read.csv("/Users/coletteberg/Dropbox/LowryLab/Pvirgatum.fourWay.RNAseq.libraryInfo.txt", sep="\t")

# format wide dataframe for filtering
counts_wide <- as.data.frame(counts_wide)
rownames(counts_wide) <- counts_wide$Gene

# transpose
counts_t <- as.data.frame(t(dplyr::select(counts_wide, -Gene)))

# paste together metadata and counts
meta_counts <- cbind(metaData, counts_t)

# only keep AP13, DAC, and F1's 
A_D <- filter(meta_counts, genotype == "AP13" | genotype == "DAC6" | genotype == "AxD_F1")

# change rownames to sample names 
rownames(A_D) <- A_D$Sample

# cleanup
AD_clean <- dplyr::select(A_D, -Library, -habitat, -genotype, -SRA.run, -SRA.study, -Sample)

# save colData for later
colData <- dplyr::select(A_D, habitat, genotype)

# transpose for next steps
AD_t <- as.data.frame(t(AD_clean))

# calc median
AD_t$median = apply(AD_t, 1, median, na.rm=T)

# only keep genes w/median expression > 3
AD_filt <- dplyr::select(filter(AD_t, median > 3), -median)

# write the data.
write.csv(AD_filt, "~/Dropbox/LowryLab/data/filtered_AP13_DAC6_RNA.csv")
write.csv(colData, "~/Dropbox/LowryLab/data/filtered_AP13_DAC6_RNA_metadata.csv")
```

# filter to just genotypes of interest 

```{r AP13 and DAC6 only}

# only keeping AP13 and DAC in MI and TX; removing DAC TX 3 bc it's a contaminant

# reformatting counts
AP_DAC <- dplyr::select(AD_filt, 
              PKLE_DAC_1,
              PKLE_AP13_1,
              PKLE_AP13_2,
              KBSM_AP13_1,
              KBSM_AP13_2,
              KBSM_DAC_1,
              KBSM_AP13_3,
              KBSM_AP13_4,
              KBSM_AP13_5,
              PKLE_DAC_4,
              KBSM_DAC_2,
              PKLE_AP13_3,
              PKLE_AP13_4,
              KBSM_DAC_3,
              PKLE_AP13_5)

# reformatting col Data
colData_AP_DAC <- filter(colData, 
       habitat == "KBSM" |
      habitat == "PKLE", 
      genotype == "DAC6" |
      genotype == "AP13") %>% slice(-9)

```


# run DESeq (make dds object)
```{r make a DESeq object}
# format for DESeq 

dds <- DESeqDataSetFromMatrix(countData = AP_DAC,
                              colData = colData_AP_DAC,
                              design =  ~ habitat + genotype + habitat:genotype)

```

# filter the DDS object
```{r filtering}
smallestGroupSize <- 2
keep <- rowSums(counts(dds) >= 10) >= smallestGroupSize
dds <- dds[keep,]
 

```

# normalize & run DESeq
```{r normalize & run}
# running deseq 
dds <- DESeq(dds)

# running VSD
vsd <- varianceStabilizingTransformation(dds)


```

# PCA plot of gene expression to vet samples
```{r PCA of genotypes}

pcaData <- plotPCA(vsd, intgroup=c("habitat", "genotype"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=habitat, shape=genotype)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme_bw() 


```

# export DE gene data 
```{r export }
# create and edit habitat contrast df
results_habitat <- as.data.frame(results(dds, name="habitat_PKLE_vs_KBSM"))
results_habitat$diffexpressed <- "NO"
results_habitat$diffexpressed[results_habitat$log2FoldChange > 1 & results_habitat$padj < 0.05] <- "UP"
results_habitat$diffexpressed[results_habitat$log2FoldChange < -1 & results_habitat$padj < 0.05] <- "DOWN"

# create and edit genotype contrast df
results_genotype <- as.data.frame(results(dds, name="genotype_DAC6_vs_AP13"))
results_genotype$diffexpressed <- "NO"
results_genotype$diffexpressed[results_genotype$log2FoldChange > 1 & results_genotype$padj < 0.05] <- "UP"
results_genotype$diffexpressed[results_genotype$log2FoldChange < -1 & results_genotype$padj < 0.05] <- "DOWN"

# create and edit full model contrast df
results_full <- as.data.frame(results(dds))
results_full$diffexpressed <- "NO"
results_full$diffexpressed[results_full$log2FoldChange > 1 & results_full$padj < 0.05] <- "UP"
results_full$diffexpressed[results_full$log2FoldChange < -1 & results_full$padj < 0.05] <- "DOWN"


write.csv(results_habitat, "~/Dropbox/LowryLab/data/DE_analyses/KBSM_PKLE_DE.csv")
write.csv(results_genotype, "~/Dropbox/LowryLab/data/DE_analyses/AP13_DAC6_DE.csv")
write.csv(results_full, "~/Dropbox/LowryLab/data/DE_analyses/full_DE.csv")

# get number of DE genes 
filter(results_full, padj < 0.05, diffexpressed == "DOWN" | diffexpressed == "UP")

```

# make volcano plot of DE genes
```{r volcano plot}

# this can be done multiple times, for each of the models 
ggplot(results_full, aes(x = log2FoldChange, y = -log10(padj), col = diffexpressed)) +
  geom_point() +
  theme_minimal() +
  scale_color_manual(values = c("NO" = "grey", "UP" = "red", "DOWN" = "blue")) +
  labs(title = "Full Model DE", x = "Log2 Fold Change", y = "-Log10 Adjusted P-value") +
  geom_vline(xintercept = c(-1, 1), col = "grey") + # Add vertical lines for LFC threshold
  geom_hline(yintercept = -log10(0.05), col = "grey") # Add horizontal line for p-value threshold

```


# upload necessary data for GO analyses 
```{r upload GO terms}
# read in the genes with GO terms from the annotation 
Pvirg_GOterms <- read.csv("~/Dropbox/LowryLab/Pavir_GOterms_multiple.csv")

# melt to make longer
longer_GOterms <- unique(melt(Pvirg_GOterms,
     id.vars=c("locusName")))

# convert to clusterProfiler format
GOterms_mult  <- as.data.frame(cbind(longer_GOterms$value,
                    longer_GOterms$locusName))

# extract names of GO terms
# Extract a named vector of all GO terms and their names
goterms <- Term(GOTERM)

# Convert to a data frame for easier use
term2name <- data.frame("GOID" = names(goterms), "term" = goterms)

```

# format DE genes for GO term analyses
```{r format for GO}

results_full$gene <- gsub(".v6.1", "", rownames(results_full))
results_genotype$gene <- gsub(".v6.1", "", rownames(results_genotype))
results_habitat$gene <- gsub(".v6.1", "", rownames(results_habitat))

```

# go analysis of DE genes 
```{r run GO analyses of DE genes}

# full model DE
fullMod <- enricher(as.vector(filter(results_full, diffexpressed!="NO")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# habitat DE
habMod <- enricher(as.vector(filter(results_habitat, diffexpressed!="NO")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# genotype DE 
genMod <- enricher(as.vector(filter(results_genotype, diffexpressed!="NO")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

# write the results
write.csv(as.data.frame(fullMod@result), "~/Dropbox/LowryLab/data/DE_analyses/GO_full_DE.csv")
write.csv(as.data.frame(habMod@result), "~/Dropbox/LowryLab/data/DE_analyses/GO_habitat_DE.csv")
write.csv(as.data.frame(genMod@result), "~/Dropbox/LowryLab/data/DE_analyses/GO_genotype_DE.csv")

dotplot(fullMod)


```


# prep for WGCNA
```{r WGCNA prep}

# extract the normalized data for WGCNA and format 
wpn_vsd <- getVarianceStabilizedData(dds)
rv_wpn <- rowVars(wpn_vsd)

# transpose
input_mat_all = t(wpn_vsd)

write.csv(wpn_vsd, "~/Dropbox/LowryLab/data/AP13_DAC6_vsdNorm.csv")

```


# pick the power level & run WGCNA for all genes; choose top hub genes  
```{r WGCNA all genes pick power and run}

allowWGCNAThreads()

#> Allowing multi-threading with up to 4 threads
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 20, by = 2))

# Call the network topology analysis function
sft = pickSoftThreshold(
  input_mat_all,             # <= Input data
  blockSize = 3000,
  powerVector = powers,
  verbose = 5
  )


# plot to select threshold
par(mfrow = c(1,2));
cex1 = 0.9;

plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)

picked_power = 5
temp_cor <- cor
cor <- WGCNA::cor

```

```{r netwk call}

# generate the network
netwk_all <- blockwiseModules(input_mat_all,       
                          power = picked_power,           
                          networkType = "signed",
                          deepSplit = 2,
                          pamRespectsDendro = F,
                          minModuleSize = 100,
                          maxBlockSize = 4000,
                          reassignThreshold = 0,
                          mergeCutHeight = 0.3,
                          saveTOMs = T,
                          saveTOMFileBase = "ER",
                          numericLabels = F,
                          verbose = 3)




# print the network so you get the number of modules
summary(as.factor(netwk_all$colors))

# calc the number of genes in each module with the cut height of 0.3

# format the data frame
names_cols <- as.data.frame(netwk_all$colors)
names_cols$gene <- rownames(names_cols)

# plot
ggplot(names_cols,
       aes(x = `netwk_all$colors`)) + 
  geom_bar()


# choose top hub gene
chooseTopHubInEachModule(
   input_mat_all, 
   netwk_all$colors, 
   power = picked_power)
```


# plot WGCNA for all genes
```{r plot WGCNA all }
 # Convert labels to colors for plotting


mergedColors_all = labels2colors(netwk_all$colors)

# using a higher cut height to merge into a more manageable number of modules
# this is old so I'm commenting it out
# mergeNet <- mergeCloseModules(input_mat_all, mergedColors_all, cutHeight=0.3)
# mergeCols <- labels2colors(mergeNet$colors)

# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  netwk_all$dendrograms[[1]],
  mergedColors_all[netwk_all$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )

module_df <- data.frame(
  gene_id = names(netwk_all$colors),
  colors = labels2colors(netwk_all$colors)
)

 
# Get Module Eigengenes per cluster
MEs0_all <- moduleEigengenes(input_mat_all, netwk_all$colors)$eigengenes

# Reorder modules so similar modules are next to each other
MEs0_all <- orderMEs(MEs0_all)
module_order_all = names(MEs0_all) %>% gsub("ME","", .)

# Add treatment names
MEs0_all$treatment = row.names(MEs0_all)

# tidy data
mME_all = MEs0_all %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order_all)
  )

# plot data 
mME_all %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")

```

# formatting WGCNA output for GO analyses
```{r formatting module output}

modules_all <- as.data.frame(netwk_all$colors)
modules_all$gene <- gsub(".v6.1", "", rownames(modules_all))
modules_all$color <- modules_all$`netwk_all$colors`

```

```{r}
mME_all %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")


plotDendroAndColors(
  netwk_all$dendrograms[[1]],
  mergedColors[netwk_all$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )
```

# GO analyses for the modules (0.05% of genes)
```{r run GO analyses of 5% modules}
greenMod <- enricher(as.vector(filter(modules_0.05, color == "green")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

brownMod <- enricher(as.vector(filter(modules_0.05, color == "brown")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

turquoiseMod <- enricher(as.vector(filter(modules_0.05, color == "turquoise")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

yellowMod <- enricher(as.vector(filter(modules_0.05, color == "yellow")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

blueMod <- enricher(as.vector(filter(modules_0.05, color == "blue")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)



dotplot(turquoiseMod) + ggtitle("turquoise module")

# habitat associated modules 
dotplot(yellowMod) + ggtitle("yellow module")
dotplot(greenMod) + ggtitle("green module")
dotplot(blueMod) + ggtitle("blue module")
```

# GO analyses for modules (all genes)
```{r}
Mod_allGenes <- enricher(as.vector(filter(modules_all, color == "green")$gene), 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=test_mult, TERM2NAME = term2name)

dotplot(Mod_allGenes)


```

