---
title: "eQTL_setup"
author: "Colette Berg"
date: "2025-07-23"
output: html_document
---

## set up script and load required packages, set up matrix eQTL settings
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(MatrixEQTL)
library(dplyr)
library(ggplot2)
library(VariantAnnotation)
library(vcfR)
library(adegenet)
library(eQTLpipeline)
library(DESeq2)
library(viridis)
library(reshape2)
library(tidyr)
library(preprocessCore)
library(fs)
library(readr)
library(clusterProfiler)
library(GO.db)


# setting up matrix eQTL settings
useModel = modelLINEAR; # modelANOVA or modelLINEAR or modelLINEAR_CROSS
output_file_name = tempfile();
errorCovariance = numeric();
```

## read in gene expression, gene and snp location, covariates files, and genotypes
```{r load expr and geno data, message = FALSE}

# reading in gene expression data for Midwest, Gulf, and Atlantic 

# MI_Midwest
geneMI_Midwest = SlicedData$new()
geneMI_Midwest$fileDelimiter = ",";    
geneMI_Midwest$LoadFile("~/Dropbox/LowryLab/data/expression/genes_MI_Midwest_normCorrected.csv")

# TX_Midwest 
geneTX_Midwest = SlicedData$new()
geneTX_Midwest$fileDelimiter = ",";    
geneTX_Midwest$LoadFile("~/Dropbox/LowryLab/data/expression/genes_TX_Midwest_normCorrected.csv")

# GxE_Midwest
geneGxE_Midwest = SlicedData$new()
geneGxE_Midwest$fileDelimiter = ",";    
geneGxE_Midwest$LoadFile("~/Dropbox/LowryLab/data/expression/genes_GxE_Midwest_normCorrected.csv")

# MI_Gulf
geneMI_Gulf = SlicedData$new()
geneMI_Gulf$fileDelimiter = ",";    
geneMI_Gulf$LoadFile("~/Dropbox/LowryLab/data/expression/genes_MI_Gulf_normCorrected.csv")

# TX_Gulf 
geneTX_Gulf = SlicedData$new()
geneTX_Gulf$fileDelimiter = ",";    
geneTX_Gulf$LoadFile("~/Dropbox/LowryLab/data/expression/genes_TX_Gulf_normCorrected.csv")

# GxE_Gulf
geneGxE_Gulf = SlicedData$new()
geneGxE_Gulf$fileDelimiter = ",";    
geneGxE_Gulf$LoadFile("~/Dropbox/LowryLab/data/expression/genes_GxE_Gulf_normCorrected.csv")

# MI_Atlantic
geneMI_Atlantic = SlicedData$new()
geneMI_Atlantic$fileDelimiter = ",";    
geneMI_Atlantic$LoadFile("~/Dropbox/LowryLab/data/expression/genes_MI_Atlantic_normCorrected.csv")

# TX_Atlantic 
geneTX_Atlantic = SlicedData$new()
geneTX_Atlantic$fileDelimiter = ",";    
geneTX_Atlantic$LoadFile("~/Dropbox/LowryLab/data/expression/genes_TX_Atlantic_normCorrected.csv")

# GxE_Atlantic
geneGxE_Atlantic = SlicedData$new()
geneGxE_Atlantic$fileDelimiter = ",";    
geneGxE_Atlantic$LoadFile("~/Dropbox/LowryLab/data/expression/genes_GxE_Atlantic_normCorrected.csv")


# gene name, chromosome, start, stop
gene_location_file_name = "~/Dropbox/LowryLab/data/AP13/AP13_gene_position.txt"

# snp name, chromosome, position (bp)
snps_location_file_name = "~/Dropbox/LowryLab/data/AP13/positions.txt"

# process smps
snpspos = read.table(snps_location_file_name, 
                     header = FALSE, 
                     stringsAsFactors = FALSE,
                     sep="\t")

# process gene names 
genepos = read.table(gene_location_file_name, 
                     header = TRUE, 
                     stringsAsFactors = FALSE,
                     sep="\t")


## Load covariates file, Midwest. in this case it's the first 5 PC's (plink)

cvrt_Midwest_GxE = SlicedData$new();
cvrt_Midwest_GxE$fileDelimiter = ",";      # the TAB character
cvrt_Midwest_GxE$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Midwest_GxE$fileSkipRows = 1;          # one row of column labels
cvrt_Midwest_GxE$fileSkipColumns = 2;       # one column of row labels
cvrt_Midwest_GxE$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_GxE.csv");

## Load covariates file, Gulf. in this case it's the first 5 PC's (plink)

cvrt_Gulf_GxE = SlicedData$new();
cvrt_Gulf_GxE$fileDelimiter = ",";      # the TAB character
cvrt_Gulf_GxE$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Gulf_GxE$fileSkipRows = 1;          # one row of column labels
cvrt_Gulf_GxE$fileSkipColumns = 2;       # one column of row labels
cvrt_Gulf_GxE$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_GxE.csv");


## Load covariates file, Atlantic. in this case it's the first 5 PC's (plink)

cvrt_Atlantic_GxE = SlicedData$new();
cvrt_Atlantic_GxE$fileDelimiter = ",";      # the TAB character
cvrt_Atlantic_GxE$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Atlantic_GxE$fileSkipRows = 1;          # one row of column labels
cvrt_Atlantic_GxE$fileSkipColumns = 2;       # one column of row labels
cvrt_Atlantic_GxE$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_GxE.csv");



## Load covariates file, Midwest. in this case it's the first 5 PC's (plink)

cvrt_Midwest_MI = SlicedData$new();
cvrt_Midwest_MI$fileDelimiter = ",";      # the TAB character
cvrt_Midwest_MI$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Midwest_MI$fileSkipRows = 1;          # one row of column labels
cvrt_Midwest_MI$fileSkipColumns = 2;       # one column of row labels
cvrt_Midwest_MI$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_MI.csv");

## Load covariates file, Gulf. in this case it's the first 5 PC's (plink)

cvrt_Gulf_MI = SlicedData$new();
cvrt_Gulf_MI$fileDelimiter = ",";      # the TAB character
cvrt_Gulf_MI$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Gulf_MI$fileSkipRows = 1;          # one row of column labels
cvrt_Gulf_MI$fileSkipColumns = 2;       # one column of row labels
cvrt_Gulf_MI$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_MI.csv");


## Load covariates file, Atlantic. in this case it's the first 5 PC's (plink)

cvrt_Atlantic_MI = SlicedData$new();
cvrt_Atlantic_MI$fileDelimiter = ",";      # the TAB character
cvrt_Atlantic_MI$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Atlantic_MI$fileSkipRows = 1;          # one row of column labels
cvrt_Atlantic_MI$fileSkipColumns = 2;       # one column of row labels
cvrt_Atlantic_MI$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_MI.csv");


## Load covariates file, Midwest. in this case it's the first 5 PC's (plink)

cvrt_Midwest_TX = SlicedData$new();
cvrt_Midwest_TX$fileDelimiter = ",";      # the TAB character
cvrt_Midwest_TX$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Midwest_TX$fileSkipRows = 1;          # one row of column labels
cvrt_Midwest_TX$fileSkipColumns = 2;       # one column of row labels
cvrt_Midwest_TX$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_TX.csv");

## Load covariates file, Gulf. in this case it's the first 5 PC's (plink)

cvrt_Gulf_TX = SlicedData$new();
cvrt_Gulf_TX$fileDelimiter = ",";      # the TAB character
cvrt_Gulf_TX$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Gulf_TX$fileSkipRows = 1;          # one row of column labels
cvrt_Gulf_TX$fileSkipColumns = 2;       # one column of row labels
cvrt_Gulf_TX$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_TX.csv");


## Load covariates file, Atlantic. in this case it's the first 5 PC's (plink)

cvrt_Atlantic_TX = SlicedData$new();
cvrt_Atlantic_TX$fileDelimiter = ",";      # the TAB character
cvrt_Atlantic_TX$fileOmitCharacters = "NA"; # denote missing values;
cvrt_Atlantic_TX$fileSkipRows = 1;          # one row of column labels
cvrt_Atlantic_TX$fileSkipColumns = 2;       # one column of row labels
cvrt_Atlantic_TX$LoadFile("~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_TX.csv");


# Midwest MI
Midwest_MI = "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_MI_allChroms.txt"
Midwest_MI_geno = SlicedData$new();
Midwest_MI_geno$fileDelimiter = "\t";       # denote missing values;
Midwest_MI_geno$fileSkipColumns = 1;    
Midwest_MI_geno$fileSkipRows = 1; # one column of row labels
Midwest_MI_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Midwest_MI_geno$LoadFile(Midwest_MI)

# Midwest GxE
Midwest_GxE = "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_GxE_allChroms.txt"
Midwest_GxE_geno = SlicedData$new();
Midwest_GxE_geno$fileDelimiter = "\t";       # denote missing values;
Midwest_GxE_geno$fileSkipColumns = 1;    
Midwest_GxE_geno$fileSkipRows = 1; # one column of row labels
Midwest_GxE_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Midwest_GxE_geno$LoadFile(Midwest_GxE)

# Midwest TX
Midwest_TX = "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_TX_allChroms.txt"
Midwest_TX_geno = SlicedData$new();
Midwest_TX_geno$fileDelimiter = "\t";       # denote missing values;
Midwest_TX_geno$fileSkipColumns = 1;    
Midwest_TX_geno$fileSkipRows = 1; # one column of row labels
Midwest_TX_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Midwest_TX_geno$LoadFile(Midwest_TX)


# Atlantic MI
Atlantic_MI = "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_MI_allChroms.txt"
Atlantic_MI_geno = SlicedData$new();
Atlantic_MI_geno$fileDelimiter = "\t";       # denote missing values;
Atlantic_MI_geno$fileSkipColumns = 1;    
Atlantic_MI_geno$fileSkipRows = 1; # one column of row labels
Atlantic_MI_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Atlantic_MI_geno$LoadFile(Atlantic_MI)

# Atlantic GxE
Atlantic_GxE = "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_GxE_allChroms.txt"
Atlantic_GxE_geno = SlicedData$new();
Atlantic_GxE_geno$fileDelimiter = "\t";       # denote missing values;
Atlantic_GxE_geno$fileSkipColumns = 1;    
Atlantic_GxE_geno$fileSkipRows = 1; # one column of row labels
Atlantic_GxE_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Atlantic_GxE_geno$LoadFile(Atlantic_GxE)

# Atlantic TX
Atlantic_TX = "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_TX_allChroms.txt"
Atlantic_TX_geno = SlicedData$new();
Atlantic_TX_geno$fileDelimiter = "\t";       # denote missing values;
Atlantic_TX_geno$fileSkipColumns = 1;    
Atlantic_TX_geno$fileSkipRows = 1; # one column of row labels
Atlantic_TX_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Atlantic_TX_geno$LoadFile(Atlantic_TX)


# Gulf MI
Gulf_MI = "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_MI_allChroms.txt"
Gulf_MI_geno = SlicedData$new();
Gulf_MI_geno$fileDelimiter = "\t";       # denote missing values;
Gulf_MI_geno$fileSkipColumns = 1;    
Gulf_MI_geno$fileSkipRows = 1; # one column of row labels
Gulf_MI_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Gulf_MI_geno$LoadFile(Gulf_MI)

# Gulf GxE
Gulf_GxE = "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_GxE_allChroms.txt"
Gulf_GxE_geno = SlicedData$new();
Gulf_GxE_geno$fileDelimiter = "\t";       # denote missing values;
Gulf_GxE_geno$fileSkipColumns = 1;    
Gulf_GxE_geno$fileSkipRows = 1; # one column of row labels
Gulf_GxE_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Gulf_GxE_geno$LoadFile(Gulf_GxE)

# Gulf TX
Gulf_TX = "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_TX_allChroms.txt"
Gulf_TX_geno = SlicedData$new();
Gulf_TX_geno$fileDelimiter = "\t";       # denote missing values;
Gulf_TX_geno$fileSkipColumns = 1;    
Gulf_TX_geno$fileSkipRows = 1; # one column of row labels
Gulf_TX_geno$fileSliceSize = 10000;      # read file in pieces of 2,000 rows
Gulf_TX_geno$LoadFile(Gulf_TX)
```

## map all eQTLs
```{r  run matrix eQTL for all populations and sites}

# eQTL_mapping GxE
me_GxE_Midwest = Matrix_eQTL_main(
  snps = Midwest_GxE_geno,
  gene = geneGxE_Midwest,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Midwest_GxE,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Midwest_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Midwest_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-6,
  pvOutputThreshold.cis = 1e-3, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)

# eQTL mapping MI
me_MI_Midwest = Matrix_eQTL_main(
  snps = Midwest_MI_geno,
  gene = geneMI_Midwest,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Midwest_MI,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Midwest_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Midwest_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-7,
  pvOutputThreshold.cis = 1e-5,
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)

# eQTL mapping TX 
me_TX_Midwest = Matrix_eQTL_main(
  snps = Midwest_TX_geno,
  gene = geneTX_Midwest,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Midwest_TX,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Midwest_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Midwest_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-7,
  pvOutputThreshold.cis = 1e-5, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)


# eQTL_mapping
me_GxE_Atlantic = Matrix_eQTL_main(
  snps = Atlantic_GxE_geno,
  gene = geneGxE_Atlantic,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Atlantic_GxE,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Atlantic_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Atlantic_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-7,
  pvOutputThreshold.cis = 1e-5, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)


me_MI_Atlantic = Matrix_eQTL_main(
  snps = Atlantic_MI_geno,
  gene = geneMI_Atlantic,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Atlantic_MI,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Atlantic_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Atlantic_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-7,
  pvOutputThreshold.cis = 1e-5, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)

me_TX_Atlantic = Matrix_eQTL_main(
  snps = Atlantic_TX_geno,
  gene = geneTX_Atlantic,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Atlantic_TX,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Atlantic_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Atlantic_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-7,
  pvOutputThreshold.cis = 1e-5, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)


# eQTL_mapping
me_GxE_Gulf = Matrix_eQTL_main(
  snps = Gulf_GxE_geno,
  gene = geneGxE_Gulf,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Gulf_GxE,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Gulf_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Gulf_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-5,
  pvOutputThreshold.cis = 1e-3, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)

me_MI_Gulf = Matrix_eQTL_main(
  snps = Gulf_MI_geno,
  gene = geneMI_Gulf,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Gulf_MI,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Gulf_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Gulf_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-5,
  pvOutputThreshold.cis = 1e-3, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)

me_TX_Gulf = Matrix_eQTL_main(
  snps = Gulf_TX_geno,
  gene = geneTX_Gulf,
  snpspos = snpspos,
  genepos = genepos,
  cisDist = 1e6,
  cvrt = cvrt_Gulf_TX,
  output_file_name = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Gulf_allChroms_normCorrected.txt",
  output_file_name.cis = "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Gulf_allChroms_cis_normCorrected.txt",
  pvOutputThreshold = 1e-5,
  pvOutputThreshold.cis = 1e-3, 
  useModel = useModel,
  errorCovariance = errorCovariance,
  verbose = TRUE,
  pvalue.hist = "qqplot",
  min.pv.by.genesnp = TRUE,
  noFDRsaveMemory = FALSE)
```

# read back in the eQTLs for post-processing and rewrite
```{r eQTL post processing}

# read in and process gene position file
genePos <- read.csv("~/Dropbox/LowryLab/data/AP13/AP13_gene_position.txt",
                    sep="\t")

# format (add gene column) for later
genePos$gene <- genePos$geneid

# read in the trans eQTL output files -- GxE
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Midwest_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Midwest_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Midwest_GxE_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")

# read in the trans eQTL output files  -- MI
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Midwest_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Midwest_allChroms_cis_normCorrected.txt", sep = "\t")
 
# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Midwest_MI_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")

# read in the trans eQTL output files -- TX
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Midwest_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Midwest_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Midwest_TX_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")





# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Atlantic_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Atlantic_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Atlantic_GxE_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")


# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Atlantic_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Atlantic_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Atlantic_MI_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")



# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Atlantic_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Atlantic_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Atlantic_TX_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")



# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Gulf_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Gulf_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Gulf_GxE_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")


# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Gulf_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Gulf_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Gulf_MI_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")


# read in the trans eQTL output files 
trans <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Gulf_allChroms_normCorrected.txt", sep = "\t")
cis <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Gulf_allChroms_cis_normCorrected.txt", sep = "\t")

# add cis label to column and merge with gene positions 
cis$type <- "cis"
cis_labeled <- merge(cis, genePos, by="gene")
trans$type <- "trans"
trans_labeled <- merge(trans, genePos, by="gene")

# put the cis and trans together! 
all_eQTLs <- rbind(cis_labeled, trans_labeled)

# now separate the SNP label for plotting
Gulf_TX_eQTLs <- all_eQTLs %>%
  separate(SNP, into = c("genome_Chrom", "genome_bp"), sep = "_")

# filter by FDR and rewrite
# filter Gulf
Gulf_TX_eQTLs_filt <- filter(Gulf_TX_eQTLs, FDR < 0.05)
Gulf_MI_eQTLs_filt <- filter(Gulf_MI_eQTLs, FDR < 0.05)
Gulf_GxE_eQTLs_filt <- filter(Gulf_GxE_eQTLs, FDR < 0.05)

# filter Atlantic
Atlantic_TX_eQTLs_filt <- filter(Atlantic_TX_eQTLs, FDR < 0.05)
Atlantic_MI_eQTLs_filt <- filter(Atlantic_MI_eQTLs, FDR < 0.05)
Atlantic_GxE_eQTLs_filt <- filter(Atlantic_GxE_eQTLs, FDR < 0.05)

# filter Atlantic
Midwest_TX_eQTLs_filt <- filter(Midwest_TX_eQTLs, FDR < 0.05)
Midwest_MI_eQTLs_filt <- filter(Midwest_MI_eQTLs, FDR < 0.05)
Midwest_GxE_eQTLs_filt <- filter(Midwest_GxE_eQTLs, FDR < 0.05)


# write filtered eQTls to save them
#write.csv(Gulf_TX_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Gulf_all_eQTLs.csv")
#write.csv(Gulf_MI_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Gulf_all_eQTLs.csv")
#write.csv(Gulf_GxE_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Gulf_all_eQTLs.csv")

# write all eQTls to save them
#write.csv(Midwest_MI_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Midwest_all_eQTLs.csv")
#write.csv(Midwest_GxE_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Midwest_all_eQTLs.csv")
#write.csv(Midwest_TX_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Midwest_all_eQTLs.csv")

# write all eQTls to save them
#write.csv(Atlantic_MI_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/MI_Atlantic_all_eQTLs.csv")
#write.csv(Atlantic_TX_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/TX_Atlantic_all_eQTLs.csv")
#write.csv(Atlantic_GxE_eQTLs_filt, "/Users/coletteberg/Dropbox/LowryLab/data/eQTL_output/GxE_Atlantic_all_eQTLs.csv")


#ggplot(filter(Gulf_TX_eQTLs, FDR < 0.2, type == "trans"),
 #      aes(x = p.value,
  #         y = FDR)) +
  #geom_point()

#nrow(filter(Atlantic_MI_eQTLs, FDR < 0.1, type =="cis"))

#hist(Gulf_TX_eQTLs$FDR)\
```

# read in genotypes and expression as csvs for subsequent processing
```{r reupload_genotypes}
# read in the genotypes for SFS
# Midwest
Midwest_MI_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_MI_allChroms.txt", sep = "\t")
Midwest_TX_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_TX_allChroms.txt", sep = "\t")
Midwest_GxE_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_GxE_allChroms.txt", sep = "\t")

#Atlantic
Atlantic_MI_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_MI_allChroms.txt", sep = "\t")
Atlantic_TX_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_TX_allChroms.txt", sep = "\t")
Atlantic_GxE_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_GxE_allChroms.txt", sep = "\t")

#Gulf
Gulf_MI_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_MI_allChroms.txt", sep = "\t")
Gulf_TX_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_TX_allChroms.txt", sep = "\t")
Gulf_GxE_genotypes <- read.csv("~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_GxE_allChroms.txt", sep = "\t")

# to explore why some genes have excessive eQTLs, re-reading in gene expression files as .csvs 
# Gulf
geneMI_Gulf <- read.csv("~/Dropbox/LowryLab/data/expression/genes_MI_Gulf_normCorrected.csv")
geneTX_Gulf <- read.csv("~/Dropbox/LowryLab/data/expression/genes_TX_Gulf_normCorrected.csv")
geneGxE_Gulf <- read.csv("~/Dropbox/LowryLab/data/expression/genes_GxE_Gulf_normCorrected.csv")
# Atlantic
geneMI_Atlantic <- read.csv("~/Dropbox/LowryLab/data/expression/genes_MI_Atlantic_normCorrected.csv")
geneTX_Atlantic <- read.csv("~/Dropbox/LowryLab/data/expression/genes_TX_Atlantic_normCorrected.csv")
geneGxE_Atlantic <- read.csv("~/Dropbox/LowryLab/data/expression/genes_GxE_Atlantic_normCorrected.csv")
# Midwest
geneMI_Midwest <- read.csv("~/Dropbox/LowryLab/data/expression/genes_MI_Midwest_normCorrected.csv")
geneTX_Midwest <- read.csv("~/Dropbox/LowryLab/data/expression/genes_TX_Midwest_normCorrected.csv")
geneGxE_Midwest <- read.csv("~/Dropbox/LowryLab/data/expression/genes_GxE_Midwest_normCorrected.csv")

```

# getting the top 10% of GxE genes to test for enrichment
```{r top GxE gene enrichment}
# first calculate the median diff in gene expression across genes 

geneGxE_Atlantic$median <- apply(dplyr::select(geneGxE_Atlantic, -X), 1, median, na.rm = TRUE)
geneGxE_Gulf$median <- apply(dplyr::select(geneGxE_Gulf, -X), 1, median, na.rm = TRUE)
geneGxE_Midwest$median <- apply(dplyr::select(geneGxE_Midwest, -X), 1, median, na.rm = TRUE)

# now get the top 10% GxE genes 
Atlantic_filtered <- geneGxE_Atlantic %>%  slice_max(median, prop = 0.1)
Gulf_filtered <- geneGxE_Gulf %>%  slice_max(median, prop = 0.1)
Midwest_filtered <- geneGxE_Midwest %>%  slice_max(median, prop = 0.1)

# now running GO terms
# the gene ontology setup is in AP13_DAC6_DESeq 
Atlantic <- enricher(Atlantic_filtered$X, 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

Gulf <- enricher(Gulf_filtered$X, 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

Midwest <- enricher(Midwest_filtered$X, 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)

dotplot(Atlantic) + ggtitle("GO term enrichment, top Atlantic GxE genes")
dotplot(Midwest) + ggtitle("GO term enrichment, top Midwest GxE genes")
dotplot(Gulf) + ggtitle("GO term enrichment, top Gulf GxE genes")
```

# scan for hotspots 

# exploring the genes that have implausibly many eQTLs 
# testing relationship between eQTL number and deviations from normal-ness 
```{r gene summary}

Gulf_MI_summary <- filter(Gulf_MI_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Gulf_TX_summary <- filter(Gulf_TX_eQTLs, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Gulf_GxE_summary <- filter(Gulf_GxE_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Atlantic_MI_summary <- filter(Atlantic_TX_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Atlantic_TX_summary <- filter(Atlantic_TX_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Atlantic_MI_summary <- filter(Atlantic_MI_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Atlantic_GxE_summary <- filter(Atlantic_GxE_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Midwest_GxE_summary <- filter(Midwest_GxE_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Midwest_MI_summary <- filter(Midwest_MI_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())

Midwest_TX_summary <- filter(Midwest_TX_eQTLs_filt, type == "trans") %>% 
  group_by(chr,
           gene) %>% 
  summarise(n = n())



# shapiro-wilk tests for normality 
# Gulf
shap_MI_Gulf <- as.data.frame(t(apply(dplyr::select(geneMI_Gulf, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_TX_Gulf <- as.data.frame(t(apply(dplyr::select(geneTX_Gulf, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_GxE_Gulf <- as.data.frame(t(apply(dplyr::select(geneGxE_Gulf, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

# Atlantic
shap_MI_Atlantic <- as.data.frame(t(apply(dplyr::select(geneMI_Atlantic, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_TX_Atlantic <- as.data.frame(t(apply(dplyr::select(geneTX_Atlantic, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_GxE_Atlantic <- as.data.frame(t(apply(dplyr::select(geneGxE_Atlantic, -X), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))


geneMI_Midwest$var <- apply(dplyr::select(geneMI_Midwest, -X), 1, var)
geneTX_Midwest$var <- apply(dplyr::select(geneTX_Midwest, -X), 1, var)
geneGxE_Midwest$var <- apply(dplyr::select(geneGxE_Midwest, -X), 1, var)

geneGxE_Midwest <- filter(geneGxE_Midwest, var > 0)
geneMI_Midwest <- filter(geneMI_Midwest, var > 0)
geneTX_Midwest <- filter(geneTX_Midwest, var > 0)


# Midwest
shap_MI_Midwest <- as.data.frame(t(apply(dplyr::select(geneMI_Midwest, -X, var), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_TX_Midwest <- as.data.frame(t(apply(dplyr::select(geneTX_Midwest, -X, var), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

shap_GxE_Midwest <- as.data.frame(t(apply(dplyr::select(geneGxE_Midwest, -X, var), 1, function(x) {
  res <- shapiro.test(x)
  c(W_statistic = res$statistic, p_value = res$p.value)
})))

# now some formatting and merging for exploration
shap_MI_Gulf$gene <- geneMI_Gulf$X
Gulf_MImerge <- merge(Gulf_MI_summary, shap_MI_Gulf)

shap_TX_Gulf$gene <- geneTX_Gulf$X
Gulf_TXmerge <- merge(Gulf_TX_summary, shap_TX_Gulf)

shap_GxE_Gulf$gene <- geneGxE_Gulf$X
Gulf_GxEmerge <- merge(Gulf_GxE_summary, shap_GxE_Gulf)

# now some formatting and merging for exploration
shap_MI_Midwest$gene <- geneMI_Midwest$X
Midwest_MImerge <- merge(Midwest_MI_summary, shap_MI_Midwest)

shap_TX_Midwest$gene <- geneTX_Midwest$X
Midwest_TXmerge <- merge(Midwest_TX_summary, shap_TX_Midwest)

shap_GxE_Midwest$gene <- geneGxE_Midwest$X
Midwest_GxEmerge <- merge(Midwest_GxE_summary, shap_GxE_Midwest)

# now some formatting and merging for exploration
shap_MI_Atlantic$gene <- geneMI_Atlantic$X
Atlantic_MImerge <- merge(Atlantic_MI_summary, shap_MI_Atlantic)

shap_TX_Atlantic$gene <- geneTX_Atlantic$X
Atlantic_TXmerge <- merge(Atlantic_TX_summary, shap_TX_Atlantic)

shap_GxE_Atlantic$gene <- geneGxE_Atlantic$X
Atlantic_GxEmerge <- merge(Atlantic_GxE_summary, shap_GxE_Atlantic)

```

##  meta-analyses of the filtered eQTLs 

# first just counting up the types of each eQTL
```{r making table of # of cis / trans eQTLs }

# Atlantic
nrow(filter(Atlantic_MI_eQTLs_filt, type == "cis"))
length(unique(filter(Atlantic_MI_eQTLs_filt, type == "cis")$gene))

nrow(filter(Atlantic_MI_eQTLs_filt, type == "trans"))
length(unique(filter(Atlantic_MI_eQTLs_filt, type == "trans")$gene))

nrow(filter(Atlantic_TX_eQTLs_filt, type == "cis"))
length(unique(filter(Atlantic_TX_eQTLs_filt, type == "cis")$gene))

nrow(filter(Atlantic_TX_eQTLs_filt, type == "trans"))
length(unique(filter(Atlantic_TX_eQTLs_filt, type == "trans")$gene))

nrow(filter(Atlantic_GxE_eQTLs_filt, type == "cis"))
length(unique(filter(Atlantic_GxE_eQTLs_filt, type == "cis")$gene))

nrow(filter(Atlantic_GxE_eQTLs_filt, type == "trans"))
length(unique(filter(Atlantic_GxE_eQTLs_filt, type == "trans")$gene))

# Gulf
nrow(filter(Gulf_MI_eQTLs_filt, type == "cis"))
length(unique(filter(Gulf_MI_eQTLs_filt, type == "cis")$gene))

nrow(filter(Gulf_MI_eQTLs_filt, type == "trans"))
length(unique(filter(Gulf_MI_eQTLs_filt, type == "trans")$gene))

nrow(filter(Gulf_TX_eQTLs_filt, type == "cis"))
length(unique(filter(Gulf_TX_eQTLs_filt, type == "cis")$gene))

nrow(filter(Gulf_TX_eQTLs_filt, type == "trans"))
length(unique(filter(Gulf_TX_eQTLs_filt, type == "trans")$gene))

nrow(filter(Gulf_GxE_eQTLs_filt, type == "cis"))
length(unique(filter(Gulf_GxE_eQTLs_filt, type == "cis")$gene))

nrow(filter(Gulf_GxE_eQTLs_filt, type == "trans"))
length(unique(filter(Gulf_GxE_eQTLs_filt, type == "trans")$gene))

# Midwest
nrow(filter(Midwest_MI_eQTLs_filt, type == "cis"))
length(unique(filter(Midwest_MI_eQTLs_filt, type == "cis")$gene))

nrow(filter(Midwest_MI_eQTLs_filt, type == "trans"))
length(unique(filter(Midwest_MI_eQTLs_filt, type == "trans")$gene))

nrow(filter(Midwest_TX_eQTLs_filt, type == "cis"))
length(unique(filter(Midwest_TX_eQTLs_filt, type == "cis")$gene))

nrow(filter(Midwest_TX_eQTLs_filt, type == "trans"))
length(unique(filter(Midwest_TX_eQTLs_filt, type == "trans")$gene))

nrow(filter(Midwest_GxE_eQTLs_filt, type == "cis"))
length(unique(filter(Midwest_GxE_eQTLs_filt, type == "cis")$gene))

nrow(filter(Midwest_GxE_eQTLs_filt, type == "trans"))
length(unique(filter(Midwest_GxE_eQTLs_filt, type == "trans")$gene))



nrow(filter(Midwest_MI_eQTLs_filt, type == "trans"))
nrow(filter(Midwest_TX_eQTLs_filt, type == "trans"))
nrow(Midwest_intersection_trans)
3821 / 44918

nrow(filter(Midwest_TX_eQTLs_filt, type == "cis"))
nrow(filter(Midwest_MI_eQTLs_filt, type == "cis"))
nrow(Midwest_intersection_cis)

7416 / 12049

```

# calculate MAFs -- all populations, all sites
```{r maf calc}

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Midwest_MI_genotypes == "0")
num_het <- rowSums(Midwest_MI_genotypes == "1")
num_alt <- rowSums(Midwest_MI_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Midwest_MI_maf <- rbind(maf_greater, maf_less)

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Midwest_TX_genotypes == "0")
num_het <- rowSums(Midwest_TX_genotypes == "1")
num_alt <- rowSums(Midwest_TX_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Midwest_TX_maf <- rbind(maf_greater, maf_less)

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Midwest_GxE_genotypes == "0")
num_het <- rowSums(Midwest_GxE_genotypes == "1")
num_alt <- rowSums(Midwest_GxE_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Midwest_GxE_maf <- rbind(maf_greater, maf_less)


# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Atlantic_MI_genotypes == "0")
num_het <- rowSums(Atlantic_MI_genotypes == "1")
num_alt <- rowSums(Atlantic_MI_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Atlantic_MI_maf <- rbind(maf_greater, maf_less)

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Atlantic_TX_genotypes == "0")
num_het <- rowSums(Atlantic_TX_genotypes == "1")
num_alt <- rowSums(Atlantic_TX_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Atlantic_TX_maf <- rbind(maf_greater, maf_less)


# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Atlantic_GxE_genotypes == "0")
num_het <- rowSums(Atlantic_GxE_genotypes == "1")
num_alt <- rowSums(Atlantic_GxE_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Atlantic_GxE_maf <- rbind(maf_greater, maf_less)


# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Gulf_MI_genotypes == "0")
num_het <- rowSums(Gulf_MI_genotypes == "1")
num_alt <- rowSums(Gulf_MI_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Gulf_MI_maf <- rbind(maf_greater, maf_less)

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Gulf_TX_genotypes == "0")
num_het <- rowSums(Gulf_TX_genotypes == "1")
num_alt <- rowSums(Gulf_TX_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Gulf_TX_maf <- rbind(maf_greater, maf_less)

# calculating MAF
# counting instances of ref / het / alt genotypes
num_ref <- rowSums(Gulf_GxE_genotypes == "0")
num_het <- rowSums(Gulf_GxE_genotypes == "1")
num_alt <- rowSums(Gulf_GxE_genotypes == "2")

# counting total alleles of alt and ref 
total_alt <- num_het + 2*num_alt
total_ref <- 2*num_ref + num_het

# adding 
total <- total_alt + total_ref

# calculating maf 
maf <- as.data.frame(total_alt / total)

# labeling for folded maf 
maf_greater <- filter(maf, `total_alt/total` >= 0.5)
maf_greater$folded <- 1 - maf_greater$`total_alt/total` 

maf_less <- filter(maf, `total_alt/total` < 0.5)
maf_less$folded <- maf_less$`total_alt/total`

Gulf_GxE_maf <- rbind(maf_greater, maf_less)

```

# now looking at intersection of eQTLs btwn habitats 
```{r intersections of eQTLs }

# intersection by habitat, cis Atlantic
Atlantic_intersection_cis <- intersect(filter(dplyr::select(Atlantic_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"),
          filter(dplyr::select(Atlantic_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"))

# intersection by habitat, trans Atlantic
Atlantic_intersection_trans <- intersect(filter(dplyr::select(Atlantic_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"),
          filter(dplyr::select(Atlantic_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"))

# intersection by habitat, cis Midwest
Midwest_intersection_cis <- intersect(filter(dplyr::select(Midwest_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"),
          filter(dplyr::select(Midwest_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"))


# intersection by habitat, trans Midwest
Midwest_intersection_trans <- intersect(filter(dplyr::select(Midwest_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"),
          filter(dplyr::select(Midwest_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"))


# intersection by habitat, cis Gulf
Gulf_intersection_cis <- intersect(filter(dplyr::select(Gulf_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"),
          filter(dplyr::select(Gulf_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis"))



# intersection by habitat, trans Gulf
Gulf_intersection_trans <- intersect(filter(dplyr::select(Gulf_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"),
          filter(dplyr::select(Gulf_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans"))



```

# make lists of those intersected eQTL and make merged eQTL files
```{r intersection docs}
# compiling and writing the list of intersection genes 
Atlantic_universal_genes <- rbind(Atlantic_intersection_cis, Atlantic_intersection_trans)
Atlantic_universal_genes$presence <- "TX_and_MI"

Gulf_universal_genes <- rbind(Gulf_intersection_cis, Gulf_intersection_trans)
Gulf_universal_genes$presence <- "TX_and_MI"

Midwest_universal_genes <- rbind(Midwest_intersection_cis, Midwest_intersection_trans)
Midwest_universal_genes$presence <- "TX_and_MI"


# now getting the set diff for each (Present ONLY in one habitat but not the others)

# doing a little setup to make this less cumbersome 

# Gulf
Gulf_MI_trans <- filter(dplyr::select(Gulf_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Gulf_MI_cis <- filter(dplyr::select(Gulf_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")
Gulf_TX_trans <- filter(dplyr::select(Gulf_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Gulf_TX_cis <- filter(dplyr::select(Gulf_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")

# Atlantic
Atlantic_MI_trans <- filter(dplyr::select(Atlantic_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Atlantic_MI_cis <- filter(dplyr::select(Atlantic_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")
Atlantic_TX_trans <- filter(dplyr::select(Atlantic_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Atlantic_TX_cis <- filter(dplyr::select(Atlantic_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")

# Midwest
Midwest_MI_trans <- filter(dplyr::select(Midwest_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Midwest_MI_cis <- filter(dplyr::select(Midwest_MI_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")
Midwest_TX_trans <- filter(dplyr::select(Midwest_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "trans")
Midwest_TX_cis <- filter(dplyr::select(Midwest_TX_eQTLs_filt, -beta, -t.stat, -p.value, -FDR), type == "cis")

# Gulf

# MI only
# now making the unique lists
MIonly_Gulf_trans <- setdiff(Gulf_MI_trans, Gulf_TX_trans)
MIonly_Gulf_cis <- setdiff(Gulf_MI_cis, Gulf_TX_cis)

# adding a label column 
MIonly_Gulf_trans$presence <- "MI_only"
MIonly_Gulf_cis$presence <- "MI_only"

# TX only
# now making the unique lists
TXonly_Gulf_trans <- setdiff(Gulf_TX_trans, Gulf_MI_trans)
TXonly_Gulf_cis <- setdiff(Gulf_TX_cis, Gulf_MI_cis)

# adding a label column 
TXonly_Gulf_trans$presence <- "TX_only"
TXonly_Gulf_cis$presence <- "TX_only"

# bind 'em
Gulf_presence_labeled <- rbind(MIonly_Gulf_trans,
      MIonly_Gulf_cis,
      TXonly_Gulf_trans,
      TXonly_Gulf_cis,
      Gulf_universal_genes)


# Atlantic

# MI only
# now making the unique lists
MIonly_Atlantic_trans <- setdiff(Atlantic_MI_trans, Atlantic_TX_trans)
MIonly_Atlantic_cis <- setdiff(Atlantic_MI_cis, Atlantic_TX_cis)

# adding a label column 
MIonly_Atlantic_trans$presence <- "MI_only"
MIonly_Atlantic_cis$presence <- "MI_only"

# TX only
# now making the unique lists
TXonly_Atlantic_trans <- setdiff(Atlantic_TX_trans, Atlantic_MI_trans)
TXonly_Atlantic_cis <- setdiff(Atlantic_TX_cis, Atlantic_MI_cis)

# adding a label column 
TXonly_Atlantic_trans$presence <- "TX_only"
TXonly_Atlantic_cis$presence <- "TX_only"

# bind 'em
Atlantic_presence_labeled <- rbind(MIonly_Atlantic_trans,
      MIonly_Atlantic_cis,
      TXonly_Atlantic_trans,
      TXonly_Atlantic_cis,
      Atlantic_universal_genes)

# Midwest

# MI only
# now making the unique lists
MIonly_Midwest_trans <- setdiff(Midwest_MI_trans, Midwest_TX_trans)
MIonly_Midwest_cis <- setdiff(Midwest_MI_cis, Midwest_TX_cis)

# adding a label column 
MIonly_Midwest_trans$presence <- "MI_only"
MIonly_Midwest_cis$presence <- "MI_only"

# TX only
# now making the unique lists
TXonly_Midwest_trans <- setdiff(Midwest_TX_trans, Midwest_MI_trans)
TXonly_Midwest_cis <- setdiff(Midwest_TX_cis, Midwest_MI_cis)

# adding a label column 
TXonly_Midwest_trans$presence <- "TX_only"
TXonly_Midwest_cis$presence <- "TX_only"

# bind 'em
Midwest_presence_labeled <- rbind(MIonly_Midwest_trans,
      MIonly_Midwest_cis,
      TXonly_Midwest_trans,
      TXonly_Midwest_cis,
      Midwest_universal_genes)

```

# merge static + dynamic eQTLs with MAFs
```{r maf merging}

# formatting to make merging easier 
Midwest_presence_labeled$posName <- paste(Midwest_presence_labeled$genome_Chrom, Midwest_presence_labeled$genome_bp, sep = "_")
Gulf_presence_labeled$posName <- paste(Gulf_presence_labeled$genome_Chrom, Gulf_presence_labeled$genome_bp, sep = "_")
Atlantic_presence_labeled$posName <- paste(Atlantic_presence_labeled$genome_Chrom, Atlantic_presence_labeled$genome_bp, sep = "_")

# formatting maf file so posName column matches 
#Gulf
Gulf_MI_maf$posName <- rownames(Gulf_MI_maf)
Gulf_TX_maf$posName <- rownames(Gulf_TX_maf)
Gulf_GxE_maf$posName <- rownames(Gulf_GxE_maf)

#Midwest
Midwest_MI_maf$posName <- rownames(Midwest_MI_maf)
Midwest_TX_maf$posName <- rownames(Midwest_TX_maf)
Midwest_GxE_maf$posName <- rownames(Midwest_GxE_maf)

#Atlantic
Atlantic_MI_maf$posName <- rownames(Atlantic_MI_maf)
Atlantic_TX_maf$posName <- rownames(Atlantic_TX_maf)
Atlantic_GxE_maf$posName <- rownames(Atlantic_GxE_maf)

# merge filtered eQTLs with mafs 
Gulf_maf <- inner_join(Gulf_presence_labeled, Gulf_MI_maf)
Midwest_maf <- inner_join(Midwest_presence_labeled, Midwest_MI_maf)
Atlantic_maf <- inner_join(Atlantic_presence_labeled, Atlantic_MI_maf)

```

# merge GxE eQTLs with MAFs
```{r merge GxE mafs}

Gulf_GxE_eQTLs_filt$posName <- paste(Gulf_GxE_eQTLs_filt$genome_Chrom, Gulf_GxE_eQTLs_filt$genome_bp, sep = "_")
Midwest_GxE_eQTLs_filt$posName <- paste(Midwest_GxE_eQTLs_filt$genome_Chrom, Midwest_GxE_eQTLs_filt$genome_bp, sep = "_")
Atlantic_GxE_eQTLs_filt$posName <- paste(Atlantic_GxE_eQTLs_filt$genome_Chrom, Atlantic_GxE_eQTLs_filt$genome_bp, sep = "_")

Gulf_GxE_maf$posName <- rownames(Gulf_GxE_maf)
Midwest_GxE_maf$posName <- rownames(Midwest_GxE_maf)
Atlantic_GxE_maf$posName <- rownames(Atlantic_GxE_maf)

# merge filtered eQTLs with mafs 
Gulf_GxE_maf_ <- inner_join(Gulf_GxE_eQTLs_filt, Gulf_GxE_maf)
Midwest_GxE_maf_ <- inner_join(Midwest_GxE_eQTLs_filt, Midwest_GxE_maf)
Atlantic_GxE_maf_ <- inner_join(Atlantic_GxE_eQTLs_filt, Atlantic_GxE_maf)

write.csv(Gulf_GxE_maf_, "~/Dropbox/LowryLab/Gulf_GxE_maf.csv")
write.csv(Atlantic_GxE_maf_, "~/Dropbox/LowryLab/Atlantic_GxE_maf.csv")
write.csv(Midwest_GxE_maf_, "~/Dropbox/LowryLab/Midwest_GxE_maf.csv")
```

# merge static + dynamic eQTLs with betas (effect size)
```{r beta merging}

# Gulf
# merge Gulf with beta stats
MI_beta <- merge(filter(Gulf_maf, presence == "MI_only"), Gulf_MI_eQTLs_filt) # merge
MI_beta$betaMI <- MI_beta$beta # relabel 
MI_beta$betaTX <- "NA" # relabel 

TX_beta <- merge(filter(Gulf_maf, presence == "TX_only"), Gulf_TX_eQTLs_filt)
TX_beta$betaMI <- "NA" # relabel 
TX_beta$betaTX <- TX_beta$beta # relabel 

betaMI_static <- merge(filter(Gulf_maf, presence == "TX_and_MI"), Gulf_MI_eQTLs_filt)
betaMI_static$betaMI <- betaMI_static$beta # relabel 

betaTX_static <- merge(filter(Gulf_maf, presence == "TX_and_MI"), Gulf_TX_eQTLs_filt)

betaMI_static$betaTX <- betaTX_static$beta

MI_TX <- dplyr::select(rbind(TX_beta,
                             MI_beta), -beta, -t.stat, -p.value, -FDR)

Gulf_presence_maf_beta <- rbind(MI_TX, dplyr::select(betaMI_static, -beta, -t.stat, -p.value, -FDR))

# Atlantic
# merge Atlantic with beta stats
MI_beta <- merge(filter(Atlantic_maf, presence == "MI_only"), Atlantic_MI_eQTLs_filt) # merge
MI_beta$betaMI <- MI_beta$beta # relabel 
MI_beta$betaTX <- "NA" # relabel 

TX_beta <- merge(filter(Atlantic_maf, presence == "TX_only"), Atlantic_TX_eQTLs_filt)
TX_beta$betaMI <- "NA" # relabel 
TX_beta$betaTX <- TX_beta$beta # relabel 

betaMI_static <- merge(filter(Atlantic_maf, presence == "TX_and_MI"), Atlantic_MI_eQTLs_filt)
betaMI_static$betaMI <- betaMI_static$beta # relabel 

betaTX_static <- merge(filter(Atlantic_maf, presence == "TX_and_MI"), Atlantic_TX_eQTLs_filt)

betaMI_static$betaTX <- betaTX_static$beta

MI_TX <- dplyr::select(rbind(TX_beta,
                             MI_beta), -beta, -t.stat, -p.value, -FDR)

Atlantic_presence_maf_beta <- rbind(MI_TX, dplyr::select(betaMI_static, -beta, -t.stat, -p.value, -FDR))

# Midwest
# merge Midwest with beta stats
MI_beta <- merge(filter(Midwest_maf, presence == "MI_only"), Midwest_MI_eQTLs_filt) # merge
MI_beta$betaMI <- MI_beta$beta # relabel 
MI_beta$betaTX <- "NA" # relabel 

TX_beta <- merge(filter(Midwest_maf, presence == "TX_only"), Midwest_TX_eQTLs_filt)
TX_beta$betaMI <- "NA" # relabel 
TX_beta$betaTX <- TX_beta$beta # relabel 

betaMI_static <- merge(filter(Midwest_maf, presence == "TX_and_MI"), Midwest_MI_eQTLs_filt)
betaMI_static$betaMI <- betaMI_static$beta # relabel 

betaTX_static <- merge(filter(Midwest_maf, presence == "TX_and_MI"), Midwest_TX_eQTLs_filt)

betaMI_static$betaTX <- betaTX_static$beta

MI_TX <- dplyr::select(rbind(TX_beta,
                             MI_beta), -beta, -t.stat, -p.value, -FDR)

Midwest_presence_maf_beta <- rbind(MI_TX, dplyr::select(betaMI_static, -beta, -t.stat, -p.value, -FDR))

# now make the betas numeric
Atlantic_presence_maf_beta$betaMI <- as.numeric(Atlantic_presence_maf_beta$betaMI)
Atlantic_presence_maf_beta$betaTX <- as.numeric(Atlantic_presence_maf_beta$betaTX)

Gulf_presence_maf_beta$betaMI <- as.numeric(Gulf_presence_maf_beta$betaMI)
Gulf_presence_maf_beta$betaTX <- as.numeric(Gulf_presence_maf_beta$betaTX)

Midwest_presence_maf_beta$betaMI <- as.numeric(Midwest_presence_maf_beta$betaMI)
Midwest_presence_maf_beta$betaTX <- as.numeric(Midwest_presence_maf_beta$betaTX)


# write for later!! 
#write.csv(Atlantic_presence_maf_beta, "~/Dropbox/LowryLab/Atlantic_presence_maf_beta.csv")
#write.csv(Gulf_presence_maf_beta, "~/Dropbox/LowryLab/Gulf_presence_maf_beta.csv")
#write.csv(Midwest_presence_maf_beta, "~/Dropbox/LowryLab/Midwest_presence_maf_beta.csv")

# writing down the eQTL
write.csv(Atlantic_presence_maf_beta, "~/Desktop/bergetal_2026/Atlantic_eQTL.csv")
write.csv(Gulf_presence_maf_beta, "~/Desktop/bergetal_2026/Gulf_eQTL.csv")
write.csv(Midwest_presence_maf_beta, "~/Desktop/bergetal_2026/Midwest_eQTL.csv")


```

# explore effect sizes, mafs
```{r plotting of MAFs and effect sizes }

mean(filter(Atlantic_GxE_maf_,
       FDR < 0.1,
       type == "trans",
       folded > 0.05)$beta)

mean(filter(Atlantic_GxE_eQTL_maf,
       FDR < 0.1,
       type == "cis",
       folded > 0.05)$beta)

mean(filter(Atlantic_GxE_eQTL_maf,
       FDR < 0.1,
       folded > 0.05)$beta)

ggplot(Atlantic_TX_eQTL_maf,
       aes(x = type,
           y = abs(beta))) +
  geom_boxplot()

ggplot(filter(Midwest_GxE_eQTL_maf),
        aes(x = folded)) +
         geom_histogram() +
  facet_wrap(~type,
             scales = "free")

ggplot(Midwest_GxE_eQTL_maf,
       aes(x = folded,
           y = abs(beta))) +
  geom_point() +
  facet_wrap(~type)

ggplot(filter(Atlantic_TX_eQTL_maf,
              folded > 0.05),
       aes(x = folded,
           y = abs(beta),
           col = type)) +
  geom_smooth(method = "lm")

ggplot(filter(Atlantic_GxE_eQTL_maf,
              folded > 0.05),
       aes(x = folded)) +
  geom_histogram() +
  facet_wrap(~type,
             scales = "free")


```

# count up static & dynamic
```{r count up static & dynamic eQTLs}
filter(Midwest_maf, folded > 0.05) %>% 
  group_by(type,
           presence) %>% 
  summarise(n = n())

filter(Gulf_maf, folded > 0.05) %>% 
  group_by(type,
           presence) %>% 
  summarise(n = n_distinct(geneid))

Midwest_presence_labeled %>% 
  group_by(type,
           presence) %>% 
  summarise(n = n_distinct(geneid))

Gulf_presence_labeled %>% 
  group_by(type,
           presence) %>% 
  summarise(n = n_distinct(geneid))

Atlantic_presence_labeled %>% 
  group_by(type,
           presence) %>% 
  summarise(n = n_distinct(geneid))


Atlantic_GxE_eQTLs_filt  %>% 
  group_by(type) %>% 
  summarise(n = n_distinct(geneid))


```

# compare numbers in N and K subgenomes
```{r compare N and K subgenomes}
Midwest_presence_labeled$population <- "Midwest"
Atlantic_presence_labeled$population <- "Atlantic"
Gulf_presence_labeled$population <- "Gulf"

allPops_labeled <- rbind(Midwest_presence_labeled,
                        Atlantic_presence_labeled,
                         Gulf_presence_labeled)

K <- filter(allPops_labeled, genome_Chrom == "Chr01K" |
              genome_Chrom == "Chr02K" |
                genome_Chrom == "Chr03K" |
                genome_Chrom == "Chr04K" |
                genome_Chrom == "Chr05K" |
                genome_Chrom == "Chr06K" |
                genome_Chrom == "Chr07K" |
                genome_Chrom == "Chr08K" |
                genome_Chrom == "Chr09K")

N <- filter(allPops_labeled, genome_Chrom == "Chr01N" |
              genome_Chrom == "Chr02N" |
                genome_Chrom == "Chr03N" |
                genome_Chrom == "Chr04N" |
                genome_Chrom == "Chr05N" |
                genome_Chrom == "Chr06N" |
                genome_Chrom == "Chr07N" |
                genome_Chrom == "Chr08N" |
                genome_Chrom == "Chr09N")

K$subgenome <- "K"
N$subgenome <- "N"            

KN_labeled <- rbind(K, N)

KN_labeled

ggplot(KN_labeled, aes(x = type,
                       fill = presence)) +
  facet_grid(~ population) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("dodgerblue",
                                "black",
                                "red2"))

ggplot(KN_labeled, aes(x = genome_Chrom,
           fill = presence)) +
  geom_bar() +
  facet_wrap(~population)

```

# get & compare means of mafs
```{r mean mafs}

# compare cis and trans eQTLs 
t.test(Midwest_MI_cisMAF$folded, Midwest_MI_transMAF$folded)
t.test(Atlantic_MI_cisMAF$folded, Atlantic_MI_transMAF$folded)
t.test(Gulf_MI_cisMAF$folded, Gulf_MI_transMAF$folded)

t.test(Midwest_TX_cisMAF$folded, Midwest_TX_transMAF$folded)
t.test(Atlantic_TX_cisMAF$folded, Atlantic_TX_transMAF$folded)
t.test(Gulf_TX_cisMAF$folded, Gulf_TX_transMAF$folded)

t.test(Midwest_MI_cisMAF$folded, Midwest_MI_transMAF$folded)
t.test(Atlantic_MI_cisMAF$folded, Atlantic_MI_transMAF$folded)
t.test(Gulf_MI_cisMAF$folded, Gulf_MI_transMAF$folded)

# compare cis eQTLs to all SNPs 

t.test(Midwest_TX_cisMAF$folded, Midwest_TX_maf$folded)
t.test(Atlantic_TX_cisMAF$folded, Atlantic_TX_maf$folded)
t.test(Gulf_TX_cisMAF$folded, Gulf_TX_maf$folded)

t.test(Midwest_MI_cisMAF$folded, Midwest_MI_maf$folded)
t.test(Atlantic_TX_cisMAF$folded, Atlantic_TX_maf$folded)
t.test(Gulf_TX_cisMAF$folded, Gulf_TX_maf$folded)

#Midwest_MI
mean(filter(Midwest_MI_maf, folded > 0.05)$folded)
mean(Midwest_MI_cisMAF$folded)
mean(Midwest_MI_transMAF$folded)


# Midwest_TX
mean(filter(Midwest_TX_maf, folded > 0.05)$folded)
mean(Midwest_TX_transMAF$folded)
mean(Midwest_TX_cisMAF$folded)

# Midwest_GxE
mean(filter(Midwest_GxE_maf, folded > 0.05)$folded)
mean(Midwest_GxE_transMAF$folded)
mean(Midwest_GxE_cisMAF$folded)

#Atlantic_MI
mean(filter(Atlantic_MI_maf, folded > 0.05)$folded)
mean(Atlantic_MI_cisMAF$folded)
mean(Atlantic_MI_transMAF$folded)

# Atlantic_TX
mean(filter(Atlantic_TX_maf, folded > 0.05)$folded)
mean(Atlantic_TX_transMAF$folded)
mean(Atlantic_TX_cisMAF$folded)

# Atlantic_GxE
mean(filter(Atlantic_GxE_maf, folded > 0.05)$folded)
mean(Atlantic_GxE_transMAF$folded)
mean(Atlantic_GxE_cisMAF$folded)

#Gulf_MI
mean(filter(Gulf_MI_maf, folded > 0.05)$folded)
mean(Gulf_MI_cisMAF$folded)
mean(Gulf_MI_transMAF$folded)

# Gulf_TX
mean(filter(Gulf_TX_maf, folded > 0.05)$folded)
mean(Gulf_TX_transMAF$folded)
mean(Gulf_TX_cisMAF$folded)

# Gulf_GxE
mean(filter(Gulf_GxE_maf, folded > 0.05)$folded)
mean(Gulf_GxE_transMAF$folded)
mean(Gulf_GxE_cisMAF$folded)



```

# summarize number of eQTLs on each chromosome
```{r summ chrom nums}

# summarize by chromosome 
chrom_summ_Gulf_MI <- filter(Gulf_MI_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

chrom_summ_Gulf_TX <- filter(Gulf_TX_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

chrom_summ_Midwest_TX <- filter(Midwest_TX_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

chrom_summ_Midwest_MI <- filter(Midwest_MI_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

chrom_summ_Atlantic_MI <- filter(Atlantic_MI_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

chrom_summ_Atlantic_TX <- filter(Atlantic_TX_eQTLs_filt,FDR < 0.05) %>% 
  group_by(genome_Chrom,
           type) %>% 
  summarise(n = n())

# label the summarize files by genome (n or k)

# Gulf MI 
K_GulfMI <- filter(chrom_summ_Gulf_MI,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_GulfMI$subgenome <- "K"

N_GulfMI <- filter(chrom_summ_Gulf_MI,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_GulfMI$subgenome <- "N"

NK_GulfMI <- rbind(K_GulfMI, N_GulfMI)

K_GulfTX <- filter(chrom_summ_Gulf_TX,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_GulfTX$subgenome <- "K"

N_GulfTX <- filter(chrom_summ_Gulf_TX,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_GulfTX$subgenome <- "N"

NK_GulfTX <- rbind(K_GulfTX, N_GulfTX)




# Atlantic MI 
K_AtlanticMI <- filter(chrom_summ_Atlantic_MI,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_AtlanticMI$subgenome <- "K"

N_AtlanticMI <- filter(chrom_summ_Atlantic_MI,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_AtlanticMI$subgenome <- "N"

NK_AtlanticMI <- rbind(K_AtlanticMI, N_AtlanticMI)

K_AtlanticTX <- filter(chrom_summ_Atlantic_TX,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_AtlanticTX$subgenome <- "K"

N_AtlanticTX <- filter(chrom_summ_Atlantic_TX,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_AtlanticTX$subgenome <- "N"

NK_AtlanticTX <- rbind(K_AtlanticTX, N_AtlanticTX)



# Midwest MI 
K_MidwestMI <- filter(chrom_summ_Midwest_MI,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_MidwestMI$subgenome <- "K"

N_MidwestMI <- filter(chrom_summ_Midwest_MI,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_MidwestMI$subgenome <- "N"

NK_MidwestMI <- rbind(K_MidwestMI, N_MidwestMI)

K_MidwestTX <- filter(chrom_summ_Midwest_TX,
       genome_Chrom == "Chr01K" |
       genome_Chrom == "Chr02K" |
      genome_Chrom == "Chr03K" |
        genome_Chrom == "Chr04K" |
        genome_Chrom == "Chr05K" |
        genome_Chrom == "Chr06K" |
        genome_Chrom == "Chr07K" |
        genome_Chrom == "Chr08K" |
        genome_Chrom == "Chr09K")

K_MidwestTX$subgenome <- "K"

N_MidwestTX <- filter(chrom_summ_Midwest_TX,
       genome_Chrom == "Chr01N" |
       genome_Chrom == "Chr02N" |
       genome_Chrom == "Chr03N" |
        genome_Chrom == "Chr04N" |
        genome_Chrom == "Chr05N" |
        genome_Chrom == "Chr06N" |
        genome_Chrom == "Chr07N" |
        genome_Chrom == "Chr08N" |
        genome_Chrom == "Chr09N")

N_MidwestTX$subgenome <- "N"

NK_MidwestTX <- rbind(K_MidwestTX, N_MidwestTX)

ggplot(NK_GulfTX,
       aes(x = genome_Chrom,
           y = n,
           fill = subgenome,
           col = subgenome)) +
  geom_bar(stat="identity",
           position = "dodge") +
  facet_wrap(~type)

ggplot(NK_GulfMI,
       aes(x = genome_Chrom,
           y = n,
           fill = subgenome,
           col = subgenome)) +
  geom_bar(stat="identity",
           position = "dodge") +
  facet_wrap(~type)

ggplot(NK_AtlanticTX,
       aes(x = genome_Chrom,
           y = n,
           fill = subgenome,
           col = subgenome)) +
  geom_bar(stat="identity",
           position = "dodge") +
  facet_wrap(~type)

ggplot(NK_MidwestTX,
       aes(x = genome_Chrom,
           y = n,
           fill = subgenome,
           col = subgenome)) +
  geom_bar(stat="identity",
           position = "dodge") +
  facet_wrap(~type)

NK_MidwestTX$pop <- "Midwest"
NK_MidwestTX$hab <- "TX"

NK_MidwestMI$pop <- "Midwest"
NK_MidwestMI$hab <- "MI"

NK_AtlanticTX$pop <- "Atlantic"
NK_AtlanticTX$hab < "TX"

NK_AtlanticMI$pop <- "Atlantic"
NK_AtlanticMI$hab <- "MI"

NK_GulfMI$pop <- "Gulf"
NK_GulfMI$hab <- "MI"

NK_GulfTX$pop <- "Gulf"
NK_GulfTX$hab <- "TX"

eQTL_all_chromSumm <- rbind(NK_MidwestTX,
                            NK_MidwestMI,
                            NK_AtlanticTX,
                            NK_AtlanticMI,
                            NK_GulfMI,
                            NK_GulfTX)

write.csv(eQTL_all_chromSumm,
          "~/Dropbox/LowryLab/data/eQTL_output/eQTL_all_chromSumm.csv")

```

## explore leaf phosphorus candidate genes and GxE 
```{r leaf phosphorus GxE}
phos <- read.csv("~/Dropbox/LowryLab/P_gxe.csv")

ggplot(phos,
       aes(y = P31,
           x = site,
           col = pop,
           group = PLANT_ID)) +
  geom_point() +
  geom_line() + 
  facet_wrap(~pop)


ggplot(phos,
       aes(x = site,
           y = P31,
           col = pop)) +
  stat_summary(aes(group = pop), fun.y = mean, geom = "path") +
  stat_summary(aes(color = pop), fun.data = mean_cl_boot, geom = "errorbar", width = 0.1) +
  stat_summary(aes(color = pop), fun.y = mean, geom = "point", size = 4) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 12)) +
  ylab("Leaf Phosphorus Content") +
  xlab("Site")
```

# overlap phosphorus candidate genes with eQTL data 
```{r P candidates}
# read in Caio's candidates from all populations
Gulf <- read.csv("~/Dropbox/LowryLab/CandidateGenes_Gulf.csv")
Atlantic <- read.csv("~/Dropbox/LowryLab/CandidateGenes_Atlantic.csv")
Midwest <- read.csv("~/Dropbox/LowryLab/CandidateGenes_Midwest.csv")

# check whether candidate genes exhibit expression GxE
intersect(Gulf_filtered$X, Gulf$locusName)
intersect(Atlantic_filtered$X, Atlantic$locusName)
intersect(Midwest_filtered$X, Midwest$locusName)

# merge the Gulf candidate genes with Gulf eQTLs 
# housekeeping -- label
Gulf$geneid <- Gulf$locusName
Atlantic$geneid <- Atlantic$locusName
Midwest$geneid <- Midwest$locusName

# merge
Gulf_P_eQTL_merge <- merge(Gulf, Gulf_maf)
Atlantic_P_eQTL_merge <- merge(Atlantic, Atlantic_maf)
Midwest_P_eQTL_merge <- merge(Midwest, Atlantic_maf)

# check how many Phosphorus candidate genes are regulated by eQTL
length(unique(Gulf_P_eQTL_merge$geneid))
length(unique(Atlantic_P_eQTL_merge$geneid))
length(unique(Midwest_P_eQTL_merge$geneid))



# combine and write the merged eQTL + P candidate genes
write.csv(Gulf_P_eQTL_merge, "~/Dropbox/LowryLab/PhosCand_eQTL_Gulf.csv")
write.csv(Atlantic_P_eQTL_merge, "~/Dropbox/LowryLab/PhosCand_eQTL_Atlantic.csv")
write.csv(Midwest_P_eQTL_merge, "~/Dropbox/LowryLab/PhosCand_eQTL_Midwest.csv")

# merge these candidate genes with the GxE eQTL
Gulf_P_GxE <- merge(Gulf, Gulf_GxE_maf_)
Atlantic_P_GxE <- merge(Atlantic, Atlantic_GxE_maf_)
Midwest_P_GxE <- merge(Midwest, Atlantic_GxE_maf_)

```

# overlap phosphorus candidate genes with hotspots and trans eQTL
```{r Phos_hotspot}
library(GenomicRanges)

# first add up how many genes each SNP regulates
Gulf_summary <- filter(Gulf_maf, type == "trans") %>% 
  group_by(genome_Chrom,
           genome_bp,
           type) %>% 
  summarise(n = n())

Midwest_summary <- filter(Midwest_maf, type == "trans") %>% 
  group_by(genome_Chrom,
           genome_bp,
           type) %>% 
  summarise(n = n())

Atlantic_summary <- filter(Atlantic_maf, type == "trans") %>% 
  group_by(genome_Chrom,
           genome_bp,
           type) %>% 
  summarise(n = n())

# now filter to just the SNPs that regulate more than 50 SNPs 
Atlantic_hot <- filter(Atlantic_summary, n > 50)
Gulf_hot <- filter(Gulf_summary, n > 50)
Midwest_hot <- filter(Midwest_summary, n > 50)

# using genomic ranges in R to check whether any candidate genes are near hotspots 
# setting the genomic ranges to within 10 KB of a hotspot
Mid_hotR <- GRanges(seqnames = Midwest_hot$genome_Chrom, ranges = IRanges(start = as.numeric(Midwest_hot$genome_bp) - 20000, end = as.numeric(Midwest_hot$genome_bp) + 2000))
names(Mid_hotR) <- c(paste0(Midwest_hot$genome_Chrom, "_", Midwest_hot$genome_bp))

Atl_hotR <- GRanges(seqnames = Atlantic_hot$genome_Chrom, ranges = IRanges(start = as.numeric(Atlantic_hot$genome_bp) - 20000, end = as.numeric(Atlantic_hot$genome_bp) + 2000))
names(Atl_hotR) <- c(paste0(Atlantic_hot$genome_Chrom, "_", Atlantic_hot$genome_bp))

Glf_hotR <- GRanges(seqnames = Gulf_hot$genome_Chrom, ranges = IRanges(start = as.numeric(Gulf_hot$genome_bp) - 20000, end = as.numeric(Gulf_hot$genome_bp) + 20000))
names(Glf_hotR) <- c(paste0(Gulf_hot$genome_Chrom, "_", Gulf_hot$genome_bp))

Glf_hotR

# now make granges objects for the candidates, and name them
Gulf_candSNPs <- GRanges(seqnames=Gulf$chr, ranges = IRanges(start = Gulf$pos, end = Gulf$pos))
names(Gulf_candSNPs) <- c(paste0(Gulf$chr, "_", Gulf$pos))

Atl_candSNPs <- GRanges(seqnames=Atlantic$chr, ranges = IRanges(start = Atlantic$pos, end = Atlantic$pos))
names(Atl_candSNPs) <- c(paste0(Atlantic$chr, "_", Atlantic$pos))

Mid_candSNPs <- GRanges(seqnames=Midwest$chr, ranges = IRanges(start = Midwest$pos, end = Midwest$pos))
names(Mid_candSNPs) <- c(paste0(Midwest$chr, "_", Midwest$pos)) # GRanges names
                    
# Find overlaps
Midwest_candHot_overlaps <- findOverlaps(Mid_candSNPs, Mid_hotR)
Atlantic_candHot_overlaps <- findOverlaps(Atl_candSNPs, Atl_hotR)
Gulf_candHot_overlaps <- findOverlaps(Gulf_candSNPs, Glf_hotR)

# building a table of hotspot overlaps
Gulf_hot_candTable <- data.frame(
  query_idx   = queryHits(Gulf_candHot_overlaps),
  subject_idx = subjectHits(Gulf_candHot_overlaps),
  query_name   = names(Gulf_candSNPs)[queryHits(Gulf_candHot_overlaps)],
  subject_name = names(Glf_hotR)[subjectHits(Gulf_candHot_overlaps)]
)




# now just do the same thing to check which candidate SNPs are in the vicinity of trans eQTL 
Mid_summaryR <- GRanges(seqnames = Midwest_summary$genome_Chrom, ranges = IRanges(start = as.numeric(Midwest_summary$genome_bp) - 2000, end = as.numeric(Midwest_summary$genome_bp) + 2000))
names(Mid_summaryR) <- c(paste0(Midwest_summary$genome_Chrom, "_", Midwest_summary$genome_bp))

Atl_summaryR <- GRanges(seqnames = Atlantic_summary$genome_Chrom, ranges = IRanges(start = as.numeric(Atlantic_summary$genome_bp) - 2000, end = as.numeric(Atlantic_summary$genome_bp) + 2000))
names(Atl_summaryR) <- c(paste0(Atlantic_summary$genome_Chrom, "_", Atlantic_summary$genome_bp))

Glf_summaryR <- GRanges(seqnames = Gulf_summary$genome_Chrom, ranges = IRanges(start = as.numeric(Gulf_summary$genome_bp) - 2000, end = as.numeric(Gulf_summary$genome_bp) + 2000))
names(Glf_summaryR) <- c(paste0(Gulf_summary$genome_Chrom, "_", Gulf_summary$genome_bp))

# candidate genes and trans eQTL overlap
Midwest_candTrans_overlaps <- findOverlaps(Mid_candSNPs, Mid_summaryR)
Atlantic_candTrans_overlaps <- findOverlaps(Atl_candSNPs, Atl_summaryR)
Gulf_candTrans_overlaps <- findOverlaps(Gulf_candSNPs, Glf_summaryR)

# now building a table of hits 
Atlantic_trans_candTable <- data.frame(
  query_idx   = queryHits(Atlantic_candTrans_overlaps),
  subject_idx = subjectHits(Atlantic_candTrans_overlaps),
  query_name   = names(Atl_candSNPs)[queryHits(Atlantic_candTrans_overlaps)],
  subject_name = names(Atl_summaryR)[subjectHits(Atlantic_candTrans_overlaps)]
)

Gulf_trans_candTable <- data.frame(
  query_idx   = queryHits(Gulf_candTrans_overlaps),
  subject_idx = subjectHits(Gulf_candTrans_overlaps),
  query_name   = names(Gulf_candSNPs)[queryHits(Gulf_candTrans_overlaps)],
  subject_name = names(Glf_summaryR)[subjectHits(Gulf_candTrans_overlaps)]
)

Midwest_trans_candTable <- data.frame(
  query_idx   = queryHits(Midwest_candTrans_overlaps),
  subject_idx = subjectHits(Midwest_candTrans_overlaps),
  query_name   = names(Mid_candSNPs)[queryHits(Midwest_candTrans_overlaps)],
  subject_name = names(Mid_summaryR)[subjectHits(Midwest_candTrans_overlaps)]
)


Atlantic_trans_candTable 
# now making a table to further sort through the candidate genes that are near trans eQTL 

# housekeeping -- changing column names for merging
Gulf_trans_candTable$posName <- Gulf_trans_candTable$subject_name
Atlantic_trans_candTable$posName <- Atlantic_trans_candTable$subject_name

write.csv(merge(Atlantic_trans_candTable, Atlantic_maf),"~/Dropbox/LowryLab/PhosCand_transOverlap.csv")

View(Atlantic_trans_candTable)

59702071 - 59702834

59701136	59702576

filter(Gulf_summary, genome_bp == 59702834)

Gulf_hot_candTable 


filter(Gulf_maf, genome_Chrom == "Chr03K", genome_bp == 59702834)$geneid

```

# now do GO analyses of the GWAS hits
```{r GWAS GO}
# read in the genes with GO terms from the annotation 
Pvirg_GOterms <- read.csv("~/Dropbox/LowryLab/data/expression/Pavir_GOterms_multiple.csv")

# melt to make longer
longer_GOterms <- unique(melt(Pvirg_GOterms,
     id.vars=c("locusName")))

# convert to clusterProfiler format
GOterms_mult  <- as.data.frame(cbind(longer_GOterms$value,
                    longer_GOterms$locusName))

# extract names of GO terms
# Extract a named vector of all GO terms and their names
goterms <- Term(GOTERM)

# Convert to a data frame for easier use
term2name <- data.frame("GOID" = names(goterms), "term" = goterms)

# Gulf GWAS results GO analysis
Texas_hotSpot_GO <- enricher(filter(Gulf_maf, genome_Chrom == "Chr03K", genome_bp == 59702834)$geneid, 
         pvalueCutoff = 0.05, 
         pAdjustMethod = "BH", 
         minGSSize = 10, 
         maxGSSize = 500, qvalueCutoff = 0.2,
         TERM2GENE=GOterms_mult, TERM2NAME = term2name)


write.csv(filter(Gulf_maf, genome_Chrom == "Chr03K", genome_bp == 59702834), "~/Dropbox/LowryLab/eQTL_GulfHotspotGenes.csv")

View(Texas_hotSpot_GO)
```

# make figure of leaf phosphorus GWAS plus eQTL
```{r LPh gwas fig}
# Atlantic
MI_Atlantic_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Atlantic_KBSM.txt", sep = "\t")
TX_Atlantic_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Atlantic_PKLE.txt", sep = "\t")
# Gulf
MI_Gulf_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Gulf_KBSM.txt", sep = "\t")
TX_Gulf_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Gulf_PKLE.txt", sep = "\t")
# Midwest
MI_Midwest_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Midwest_KBSM.txt", sep = "\t")
TX_Midwest_PhGWAS <- read.csv("/Users/coletteberg/Dropbox/LowryLab/SNPs_P31_Midwest_PKLE.txt", sep = "\t")

ggplot(filter(TX_Atlantic_PhGWAS, chrom != 19), aes(x = pos,
           y = -log10(p.value),
           col = regulated,
           shape = near_transeQTL)) +
  facet_wrap(~chrom, nrow = 1, scales = "free_x") +
  geom_point(size = 0.5) +
  geom_hline(yintercept = 5.3,
             linetype = "dashed") +
  scale_color_viridis(discrete = TRUE)

# making the chrom label into a factor to re-label
MI_Atlantic_PhGWAS$chrom <- as.factor(MI_Atlantic_PhGWAS$chrom)
TX_Atlantic_PhGWAS$chrom <- as.factor(TX_Atlantic_PhGWAS$chrom)

MI_Gulf_PhGWAS$chrom <- as.factor(MI_Gulf_PhGWAS$chrom)
TX_Gulf_PhGWAS$chrom <- as.factor(TX_Gulf_PhGWAS$chrom)

MI_Midwest_PhGWAS$chrom <- as.factor(MI_Midwest_PhGWAS$chrom)
TX_Midwest_PhGWAS$chrom <- as.factor(TX_Midwest_PhGWAS$chrom)


       
```

# candidate gene investigations 
# first, reading in the gff so I can get the locations of each gene 
```{r}
gff <- read.csv("~/Downloads/")
filter(Midwest_presence_labeled, gene == "Pavir.1KG005400")

gene_locs <- read.csv("/Users/coletteberg/Downloads/download.20260406.103402/Phytozome/PhytozomeV13/Pvirgatumvar_AP13HAP1/v6.1/annotation/Pvirgatumvar_AP13HAP1_772_v6.1.gene.gff3", sep = "\t", skip=6, header = FALSE)

filt <-  filter(gene_locs, V3 == "gene")

write.csv(filt, "~/Dropbox/LowryLab/AP13_geneLocs.csv")

# using the list of fitness candidates from Lovell et al. 2021 

filter(Atlantic_presence_labeled, gene == "Pavir.1KG005300")


```
