---
title: "eQTL_setup"
author: "Colette Berg"
date: "2025-07-23"
output: html_document
---

## set up script and load required packages 
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(DESeq2)
library(viridis)
library(reshape2)
library(tidyr)
library(preprocessCore)
library(fs)
library(readr)



base.dir = find.package("MatrixEQTL")

```

# read in gene expression data, median filter, normalize, and split by population
```{r gene expression read-in}

# read in the raw data
expressionData <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/expression/switchgrass_diversityPanel_rawCounts.txt", sep="\t")

# trying this simple norm / transform from the Saunak Sen lab 

# calculate median gene expression for filtering
expressionData$row_median <- as.numeric(apply(dplyr::select(expressionData, -GeneID), 1, median, na.rm=TRUE))

# filter by median expression 
exp_filt <- filter(expressionData, row_median > 3)

# label rows 
rownames(exp_filt) <- exp_filt$GeneID

# remove median & sum columns 
exp_filt <- dplyr::select(exp_filt, -GeneID, -row_median)


# dividing every value by the rowSum to standardize by library size 
size_norm <- apply(exp_filt, 1, function(i) log2((i/sum(i) * 1000000)+1))


# now do the INT transformation 
# qnorm((rank(size_norm,na.last="keep")-0.5)/sum(!is.na(size_norm )))

as.data.frame(t(size_norm))

normalized_matrix
```



# transpose expression data, add SRR ID's & pop info
```{r transposing & adding SRR id's to the normalized data}
# read in updated metadata sheet 
metadat <- read.csv("~/Dropbox/LowryLab/data/expression/RNA_IDs_SRR.csv")

# put the expression and the metadata together
exp_plus_SRR <- cbind(metadat, size_norm)

# filter out NAs 
full_filt <- filter(exp_plus_SRR, SRR_ID != "NA")

# split by habitat
M <- filter(full_filt, Habitat == "M")
P <- filter(full_filt, Habitat == "P")

# filter to the indvs. present in both habitats (for subtracting for GxE)
M_filt <- filter(M, SRR_ID %in% intersect(M$SRR_ID, P$SRR_ID))
P_filt <- filter(P, SRR_ID %in% intersect(M$SRR_ID, P$SRR_ID))

# sort 
M_sort <- arrange(M_filt, SRR_ID)
P_sort <- arrange(P_filt, SRR_ID)

# now they are in the same order so you can subtract the difference 
diff_dat <- (dplyr::select(M_sort, -RNA_ID, -SRR_ID, -Habitat, -Population) -
             dplyr::select(P_sort, -RNA_ID, -SRR_ID, -Habitat, -Population))

# add back in the metadata columns
GxE <- cbind(dplyr::select(M_sort, SRR_ID, Habitat, Population), diff_dat)

# now filter by population
GxE_Midwest_exp <- filter(GxE, Population == "Midwest")
GxE_Atlantic_exp <- filter(GxE, Population == "Atlantic")
GxE_Gulf_exp <- filter(GxE, Population == "Gulf")

# filter the MI & TX by population as well 

# MI
MI_Midwest_exp <- filter(M, Population == "Midwest")
MI_Atlantic_exp <- filter(M, Population == "Atlantic")
MI_Gulf_exp <- filter(M, Population == "Gulf")

# TX
TX_Midwest_exp <- filter(P, Population == "Midwest")
TX_Atlantic_exp <- filter(P, Population == "Atlantic")
TX_Gulf_exp <- filter(P, Population == "Gulf")


# now perform quantile normalization by pop 

#Gulf
qNorm_TX_Gulf <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(TX_Gulf_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_MI_Gulf <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(MI_Gulf_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_GxE_Gulf <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(GxE_Gulf_exp, -SRR_ID, -Habitat, -Population)))))

# Atlantic
qNorm_TX_Atlantic <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(TX_Atlantic_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_MI_Atlantic <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(MI_Atlantic_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_GxE_Atlantic <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(GxE_Atlantic_exp, -SRR_ID, -Habitat, -Population)))))

# Midwest
qNorm_TX_Midwest <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(TX_Midwest_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_MI_Midwest <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(MI_Midwest_exp, -RNA_ID, -SRR_ID, -Habitat, -Population)))))
qNorm_GxE_Midwest <- as.data.frame(normalize.quantiles(as.matrix(t(dplyr::select(GxE_Midwest_exp, -SRR_ID, -Habitat, -Population)))))

# now do some formatting 
# Gulf
colnames(qNorm_GxE_Gulf) <- rownames(GxE_Gulf_exp)
colnames(qNorm_MI_Gulf) <- rownames(MI_Gulf_exp)
colnames(qNorm_TX_Gulf) <- rownames(TX_Gulf_exp)

# Atlantic
colnames(qNorm_GxE_Atlantic) <- rownames(GxE_Atlantic_exp)
colnames(qNorm_MI_Atlantic) <- rownames(MI_Atlantic_exp)
colnames(qNorm_TX_Atlantic) <- rownames(TX_Atlantic_exp)

# Midwest
colnames(qNorm_GxE_Midwest) <- rownames(GxE_Midwest_exp)
colnames(qNorm_MI_Midwest) <- rownames(MI_Midwest_exp)
colnames(qNorm_TX_Midwest) <- rownames(TX_Midwest_exp)

# rowname time 
#Guld
rownames(qNorm_MI_Gulf) <- colnames(dplyr::select(MI_Gulf_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_TX_Gulf) <- colnames(dplyr::select(TX_Gulf_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_GxE_Gulf) <- colnames(dplyr::select(GxE_Gulf_exp, -SRR_ID, -Habitat, -Population))

#Atlantic
rownames(qNorm_MI_Atlantic) <- colnames(dplyr::select(MI_Atlantic_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_TX_Atlantic) <- colnames(dplyr::select(TX_Atlantic_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_GxE_Atlantic) <- colnames(dplyr::select(GxE_Atlantic_exp, -SRR_ID, -Habitat, -Population))

# Midwest
rownames(qNorm_MI_Midwest) <- colnames(dplyr::select(MI_Midwest_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_TX_Midwest) <- colnames(dplyr::select(TX_Midwest_exp, -RNA_ID, -SRR_ID, -Habitat, -Population))
rownames(qNorm_GxE_Midwest) <- colnames(dplyr::select(GxE_Midwest_exp, -SRR_ID, -Habitat, -Population))

# now add in important columns
# Gulf
MI_Gulf_qexp <- cbind(dplyr::select(MI_Gulf_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_MI_Gulf)))
TX_Gulf_qexp <- cbind(dplyr::select(TX_Gulf_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_TX_Gulf)))
GxE_Gulf_qexp <- cbind(dplyr::select(GxE_Gulf_exp, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_GxE_Gulf)))

# Atlantic
MI_Atlantic_qexp <- cbind(dplyr::select(MI_Atlantic_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_MI_Atlantic)))
TX_Atlantic_qexp <- cbind(dplyr::select(TX_Atlantic_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_TX_Atlantic)))
GxE_Atlantic_qexp <- cbind(dplyr::select(GxE_Atlantic_exp, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_GxE_Atlantic)))

# Midwest
MI_Midwest_qexp <- cbind(dplyr::select(MI_Midwest_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_MI_Midwest)))
TX_Midwest_qexp <- cbind(dplyr::select(TX_Midwest_exp, RNA_ID, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_TX_Midwest)))
GxE_Midwest_qexp <- cbind(dplyr::select(GxE_Midwest_exp, SRR_ID, Habitat, Population), as.data.frame(t(qNorm_GxE_Midwest)))
```

# order the expression files so they match the genotype order
```{r ordering expression files}

# loading in the ordered SRR ID's for Midwest, Gulf, and Atlantic
Midwest_SRR <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/genotypes/Midwest_allChroms.eigenvec", sep="\t", row.names = NULL)
Gulf_SRR <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/genotypes/Gulf_allChroms.eigenvec", sep="\t", row.names = NULL)
Atlantic_SRR <- read.csv("/Users/coletteberg/Dropbox/LowryLab/data/genotypes/Atlantic_allChroms.eigenvec", sep="\t", row.names = NULL)

# now get the expression files into the same order as the genotype files 
custom_order_Midwest <- Midwest_SRR$X.IID
custom_order_Gulf <- Gulf_SRR$X.IID
custom_order_Atlantic <- Atlantic_SRR$X.IID

# order the GxE dfs
GxE_Midwest_ordered <- GxE_Midwest_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Midwest)) %>%
  arrange(GxE_Midwest_exp)

GxE_Gulf_ordered <- GxE_Gulf_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Gulf)) %>%
  arrange(GxE_Gulf_exp)

GxE_Atlantic_ordered <- GxE_Atlantic_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Atlantic)) %>%
  arrange(GxE_Atlantic_exp)

# order the MI dfs
MI_Midwest_ordered <- MI_Midwest_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Midwest)) %>%
  arrange(MI_Midwest_exp)

MI_Gulf_ordered <- MI_Gulf_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Gulf)) %>%
  arrange(MI_Gulf_exp)

MI_Atlantic_ordered <- MI_Atlantic_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Atlantic)) %>%
  arrange(MI_Atlantic_exp)

# order the TX dfs
TX_Midwest_ordered <- TX_Midwest_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Midwest)) %>%
  arrange(TX_Midwest_exp)

TX_Gulf_ordered <- TX_Gulf_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Gulf)) %>%
  arrange(TX_Gulf_exp)

TX_Atlantic_ordered <- TX_Atlantic_qexp %>%
  mutate(SRR_ID = factor(SRR_ID, levels = custom_order_Atlantic)) %>%
  arrange(TX_Atlantic_exp)

```



# format the ordered expression files 
```{r formatting expression files }

# add rownames, remove unnecessary columns, transpose 
# GxE, all three pops
rownames(GxE_Midwest_ordered) <- GxE_Midwest_ordered$SRR_ID
GxE_Midwest_ordered_formatted <- dplyr::select(GxE_Midwest_ordered, -SRR_ID, -Habitat, -Population)
t_GxE_Midwest <- as.data.frame(t(GxE_Midwest_ordered_formatted))

rownames(GxE_Atlantic_ordered) <- GxE_Atlantic_ordered$SRR_ID
GxE_Atlantic_ordered_formatted <- dplyr::select(GxE_Atlantic_ordered, -SRR_ID, -Habitat, -Population)
t_GxE_Atlantic <- as.data.frame(t(GxE_Atlantic_ordered_formatted))

rownames(GxE_Gulf_ordered) <- GxE_Gulf_ordered$SRR_ID
GxE_Gulf_ordered_formatted <- dplyr::select(GxE_Gulf_ordered, -SRR_ID, -Habitat, -Population)
t_GxE_Gulf <- as.data.frame(t(GxE_Gulf_ordered_formatted))

# MI, all three pops
rownames(MI_Midwest_ordered) <- MI_Midwest_ordered$SRR_ID
MI_Midwest_ordered_formatted <- dplyr::select(MI_Midwest_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_MI_Midwest <- as.data.frame(t(MI_Midwest_ordered_formatted))

rownames(MI_Atlantic_ordered) <- MI_Atlantic_ordered$SRR_ID
MI_Atlantic_ordered_formatted <- dplyr::select(MI_Atlantic_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_MI_Atlantic <- as.data.frame(t(MI_Atlantic_ordered_formatted))

rownames(MI_Gulf_ordered) <- MI_Gulf_ordered$SRR_ID
MI_Gulf_ordered_formatted <- dplyr::select(MI_Gulf_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_MI_Gulf <- as.data.frame(t(MI_Gulf_ordered_formatted))

# TX, all three pops
rownames(TX_Midwest_ordered) <- TX_Midwest_ordered$SRR_ID
TX_Midwest_ordered_formatted <- dplyr::select(TX_Midwest_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_TX_Midwest <- as.data.frame(t(TX_Midwest_ordered_formatted))

rownames(TX_Atlantic_ordered) <- TX_Atlantic_ordered$SRR_ID
TX_Atlantic_ordered_formatted <- dplyr::select(TX_Atlantic_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_TX_Atlantic <- as.data.frame(t(TX_Atlantic_ordered_formatted))

rownames(TX_Gulf_ordered) <- TX_Gulf_ordered$SRR_ID
TX_Gulf_ordered_formatted <- dplyr::select(TX_Gulf_ordered, -SRR_ID, -Habitat, -Population, -RNA_ID)
t_TX_Gulf <- as.data.frame(t(TX_Gulf_ordered_formatted))
```

# write expression files. filter GxE by median difference, leave the rest as is 
```{r write expression files}

# filter and write GxE files 
# GxE Midwest
t_GxE_Midwest$row_median <- apply(t_GxE_Midwest, 1, median, na.rm=TRUE)
GxE_Midwest_filt <- filter(t_GxE_Midwest, abs(row_median) > 0.5)
write.csv(dplyr::select(GxE_Midwest_filt, -row_median), "~/Dropbox/LowryLab/data/expression/genes_GxE_Midwest_quantNorm.csv")

# GxE Atlantic
t_GxE_Atlantic$row_median <- apply(t_GxE_Atlantic, 1, median, na.rm=TRUE)
GxE_Atlantic_filt <- filter(t_GxE_Atlantic, abs(row_median) > 0.5)
write.csv(dplyr::select(GxE_Atlantic_filt, -row_median), "~/Dropbox/LowryLab/data/expression/genes_GxE_Atlantic_quantNorm.csv")

# GxE Gulf
t_GxE_Gulf$row_median <- apply(t_GxE_Gulf, 1, median, na.rm=TRUE)
GxE_Gulf_filt <- filter(t_GxE_Gulf, abs(row_median) > 0.5)
write.csv(dplyr::select(GxE_Gulf_filt, -row_median), "~/Dropbox/LowryLab/data/expression/genes_GxE_Gulf_quantNorm.csv")


# write MI files
write.csv(t_MI_Midwest, "~/Dropbox/LowryLab/data/expression/genes_MI_Midwest_quantNorm.csv")
write.csv(t_MI_Atlantic, "~/Dropbox/LowryLab/data/expression/genes_MI_Atlantic_quantNorm.csv")
write.csv(t_MI_Gulf, "~/Dropbox/LowryLab/data/expression/genes_MI_Gulf_quantNorm.csv")


# write TX files 
write.csv(t_TX_Midwest, "~/Dropbox/LowryLab/data/expression/genes_TX_Midwest_quantNorm.csv")
write.csv(t_TX_Atlantic, "~/Dropbox/LowryLab/data/expression/genes_TX_Atlantic_quantNorm.csv")
write.csv(t_TX_Gulf, "~/Dropbox/LowryLab/data/expression/genes_TX_Gulf_quantNorm.csv")
```

# we need to format distinct covariates files for each habitat x population combo.
```{r covariates formatting}

# read in each cov file
MidwestCov <- read.csv("~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest.csv")
AtlanticCov <- read.csv("~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic.csv")
GulfCov <- read.csv("~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf.csv")

# GxE Covariate files 
gxe_cvrt_Midwest <- dplyr::select(MidwestCov, colnames(dplyr::select(t_GxE_Midwest, -row_median)))
gxe_cvrt_Atlantic <- dplyr::select(AtlanticCov, colnames(dplyr::select(t_GxE_Atlantic, -row_median)))
gxe_cvrt_Gulf <- dplyr::select(GulfCov, colnames(dplyr::select(t_GxE_Gulf, -row_median)))

gxe_cvrt_Midwest_1 <- cbind(dplyr::select(MidwestCov, X), gxe_cvrt_Midwest)
gxe_cvrt_Atlantic_1 <- cbind(dplyr::select(AtlanticCov, X), gxe_cvrt_Atlantic)
gxe_cvrt_Gulf_1 <- cbind(dplyr::select(GulfCov, X), gxe_cvrt_Gulf)

write.csv(gxe_cvrt_Midwest_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_GxE.csv")
write.csv(gxe_cvrt_Atlantic_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_GxE.csv")
write.csv(gxe_cvrt_Gulf_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_GxE.csv")

# MI Covariate files 
MI_cvrt_Midwest <- dplyr::select(MidwestCov, colnames(t_MI_Midwest))
MI_cvrt_Atlantic <- dplyr::select(AtlanticCov, colnames(t_MI_Atlantic))
MI_cvrt_Gulf <- dplyr::select(GulfCov, colnames(t_MI_Gulf))

MI_cvrt_Midwest_1 <- cbind(dplyr::select(MidwestCov, X), MI_cvrt_Midwest)
MI_cvrt_Atlantic_1 <- cbind(dplyr::select(AtlanticCov, X), MI_cvrt_Atlantic)
MI_cvrt_Gulf_1 <- cbind(dplyr::select(GulfCov, X), MI_cvrt_Gulf)

write.csv(MI_cvrt_Midwest_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_MI.csv")
write.csv(MI_cvrt_Atlantic_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_MI.csv")
write.csv(MI_cvrt_Gulf_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_MI.csv")

# TX Covariate files 
TX_cvrt_Midwest <- dplyr::select(MidwestCov, colnames(t_TX_Midwest))
TX_cvrt_Atlantic <- dplyr::select(AtlanticCov, colnames(t_TX_Atlantic))
TX_cvrt_Gulf <- dplyr::select(GulfCov, colnames(t_TX_Gulf))

TX_cvrt_Midwest_1 <- cbind(dplyr::select(MidwestCov, X), TX_cvrt_Midwest)
TX_cvrt_Atlantic_1 <- cbind(dplyr::select(AtlanticCov, X), TX_cvrt_Atlantic)
TX_cvrt_Gulf_1 <- cbind(dplyr::select(GulfCov, X), TX_cvrt_Gulf)

write.csv(TX_cvrt_Midwest_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Midwest_TX.csv")
write.csv(TX_cvrt_Atlantic_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Atlantic_TX.csv")
write.csv(TX_cvrt_Gulf_1, "~/Dropbox/lowryLab/data/expression/cov_allChroms_Gulf_TX.csv")

```

# now we need to read in a distinct set of genotypes for each population x hab combo, because of the different indv. numbers in each.
# testing out a faster way to read in 18 files 
```{r read-in and format}

# read them in all at once
Midwest_genotype_files <- dir_ls(path = "~/Dropbox/LowryLab/data/genotypes/Midwest/", regexp = "\\.txt$")
all_genotypes_Midwest <- read_tsv(Midwest_genotype_files, col_names = FALSE)

Atlantic_genotype_files <- dir_ls(path = "~/Dropbox/LowryLab/data/genotypes/Atlantic/", regexp = "\\.txt$")
all_genotypes_Atlantic <- read_tsv(Atlantic_genotype_files, col_names = FALSE)

Gulf_genotype_files <- dir_ls(path = "~/Dropbox/LowryLab/data/genotypes/Gulf/", regexp = "\\.txt$")
all_genotypes_Gulf <- read_tsv(Gulf_genotype_files, col_names = FALSE)

# change from tibble to dataframe
all_genotypes_Midwest <- as.data.frame(all_genotypes_Midwest)
all_genotypes_Atlantic <- as.data.frame(all_genotypes_Atlantic)
all_genotypes_Gulf <- as.data.frame(all_genotypes_Gulf)

# do some formatting  - add rownames 
rownames(all_genotypes_Midwest) <- all_genotypes_Midwest$X1
rownames(all_genotypes_Atlantic) <- all_genotypes_Atlantic$X1
rownames(all_genotypes_Gulf) <- all_genotypes_Gulf$X1

# then remove the column bc now it's a row name 
all_genotypes_Midwest <- dplyr::select(all_genotypes_Midwest, -X1)
all_genotypes_Atlantic <- dplyr::select(all_genotypes_Atlantic, -X1)
all_genotypes_Gulf <- dplyr::select(all_genotypes_Gulf, -X1)

# add in col names 
colnames(all_genotypes_Midwest) <- Midwest_SRR$X.IID
colnames(all_genotypes_Atlantic) <- Atlantic_SRR$X.IID
colnames(all_genotypes_Gulf) <- Gulf_SRR$X.IID



# now write genotypes, but including the individuals that are present in each pop-habitat combo.
# now these will be genotype files that have ALL chromosomes. 

# write GxE files 
write.table(dplyr::select(all_genotypes_Gulf, colnames(dplyr::select(GxE_Gulf_filt, -row_median))), "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_GxE_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Midwest, colnames(dplyr::select(GxE_Midwest_filt, -row_median))), "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_GxE_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Atlantic, colnames(dplyr::select(GxE_Atlantic_filt, -row_median))), "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_GxE_allChroms.txt", sep = "\t")

# write MI files 
write.table(dplyr::select(all_genotypes_Gulf, colnames(t_MI_Gulf)), "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_MI_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Midwest, colnames(t_MI_Midwest)), "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_MI_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Atlantic, colnames(t_MI_Atlantic)), "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_MI_allChroms.txt", sep = "\t")

# write TX files 
write.table(dplyr::select(all_genotypes_Gulf, colnames(t_TX_Gulf)), "~/Dropbox/lowryLab/data/genotypes/Gulf/Gulf_TX_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Midwest, colnames(t_TX_Midwest)), "~/Dropbox/lowryLab/data/genotypes/Midwest/Midwest_TX_allChroms.txt", sep = "\t")
write.table(dplyr::select(all_genotypes_Atlantic, colnames(t_TX_Atlantic)), "~/Dropbox/lowryLab/data/genotypes/Atlantic/Atlantic_TX_allChroms.txt", sep = "\t")
```



